### 2016-03-28 paper/paper.tex
Start

\def\year{2016}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai16}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\usepackage{cleveref}
\usepackage{complexity}
\usepackage[ruled,vlined]{algorithm2e} % algorithms
\usepackage{longtable}
\usepackage{booktabs}                  % less ugly tables
\usepackage{tikz}                      % For pretty pictures
\usepackage{afterpage}

\usetikzlibrary{arrows, shadows, calc, positioning, decorations, decorations.pathmorphing,
decorations.pathreplacing, patterns, tikzmark, fit}

% labels in algorithms
\newcommand{\mclabel}[1]{\label{line:mc:#1}}
\newcommand{\mcline}[1]{line~\ref{line:mc:#1}}
\newcommand{\mclinerange}[2]{lines~\ref{line:mc:#1} to~\ref{line:mc:#2}}

\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
/Author (Ciaran McCreesh, Patrick Prosser)}
\setcounter{secnumdepth}{0}

% common names of things
\newcommand{\Cbest}{C^\star}
\newcommand{\bounds}{\mathit{bounds}}
\newcommand{\order}{\mathit{order}}
\newcommand{\colour}{\mathit{colour}}
\newcommand{\uncoloured}{\mathit{uncoloured}}
\newcommand{\colourable}{\mathit{colourable}}
\newcommand{\expand}{\FuncSty{expand}}
\newcommand{\colourOrder}{\FuncSty{colourOrder}}
\newcommand{\vrej}{v_{\mathit{rej}}}
\newcommand{\kwunset}{\KwSty{unset}}
\newcommand{\vertexset}{V}
\newcommand{\neighbourhood}{N}

% some pretty colours
\definecolor{safelightblue}{rgb}{0.65098, 0.807843, 0.890196}
\definecolor{safedarkblue}{rgb}{0.121569, 0.470588, 0.705882}
\definecolor{safelightgreen}{rgb}{0.698039, 0.87451, 0.541176}
\definecolor{safedarkgreen}{rgb}{0.2, 0.627451, 0.172549}
\definecolor{safelightorange}{rgb}{0.996078, 0.901961, 0.807843}
\definecolor{safedarkorange}{rgb}{0.54902, 0.176471, 0.015686}
\definecolor{safemediumorange}{rgb}{0.74902, 0.505882, 0.490196}
\definecolor{safeverydarkblue}{rgb}{0.007843, 0.219608, 0.345098}
\definecolor{safenearlywhite}{rgb}{0.9, 0.9, 0.9}
\definecolor{safelightgrey}{rgb}{0.7, 0.7, 0.7}

\begin{document}
% The file aaai.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%
\title{Finding Maximum $k$-Cliques Faster using Lazy Global Domination}
\author{
    Ciaran McCreesh%
    \thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/K503058/1]}
    \and
    Patrick Prosser
    \\ University of Glasgow, Glasgow, Scotland
}
\maketitle
\begin{abstract}
    A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in
    this set. A $k$-clique relaxes this requirement, requiring vertices to be within a distance $k$
    of each other, rather than directly adjacent. In theory, a maximum clique algorithm can easily
    be adapted to solve the maximum $k$-clique problem. We use a state of the art maximum clique
    algorithm to show that this is feasible in practice, and introduce a lazy global domination rule
    which sometimes vastly reduces the search space. We include experimental results for a range of
    real-world and benchmark graphs, and a detailed look at random graphs.
\end{abstract}

\section{Introduction}

\begin{figure}[b] % first page please
    \centering
    \begin{tikzpicture}
        % \input{figure-k-clique-1}
    \end{tikzpicture}\hspace{4em}\begin{tikzpicture}
        % \input{figure-k-clique-2}
    \end{tikzpicture}

    \caption{On the left, a graph, with its unique maximum clique $\{1, 2, 5, 8\}$ of size 4
        highlighted. On the right, the same graph, with a maximum $2$-clique $\{1, 2, 3, 4, 5, 6, 8\}$
        of size 7 highlighted. This is not a $2$-club, since the only path of length 2 between vertices 3 and 6
        goes through vertex 7. A $3$-clique covers the entire graph.}
    \label{figure:k-cliques}
\end{figure}

A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in the
set. Finding a clique of maximum size in a graph is one of the basic \NP-hard
problems \cite{Garey:1990}; applications include geometry, coding theory, computer vision and
bioinformatics \cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may
be too strong a requirement. A $k$-clique (or sometimes $n$-clique or $s$-clique) is a relaxed form
of clique, where instead of requiring each pair of vertices to be directly adjacent, we only require
that they be connected by a path of length at most $k$ \cite{Luce:1950}. Thus a $1$-clique \emph{is} a
clique, a $2$-clique may be thought of as ``a group of people, all of whom either know each other or
have a mutual acquaintance'', and so on. We illustrate this in \cref{figure:k-cliques}.
Determining the size of a maximum $k$-clique is \NP-hard for any fixed $k$ \cite{Bourjolly:2002}.

A related relaxation is a $k$-club, which tightens the requirement of a $k$-clique as follows
\cite{Mokken:1979}. In a $k$-clique, each pair of vertices is connected by a path of length at most
$k$, but that path may use any vertices in the original graph. In a $k$-club, each pair of vertices
must be connected by a path of length at most $k$ using only vertices that are also in the club.
Thus the $2$-clique in \cref{figure:k-cliques} is \emph{not} a $2$-club (obviously, every $k$-club
is a $k$-clique).

A recent survey by Shahinpour and Butenko discusses algorithms and results for $k$-clique and
$k$-club problems \cite{Shahinpour:2013}.  We adopt their notation of $\tilde{\omega}_k$ for the
size of a maximum $k$-clique; the use of $\omega$ for the size of a maximum clique is standard.
They note that ``unlike the maximum clique problem, the maximum $s$-clique problem has not been the
subject of extensive research and we are not aware of any computational results for this problem to
date''. This is in contrast to the $k$-club problem, for which a wide range of computational results
are available
\cite{Bourjolly:2000,Bourjolly:2002,Mahdavi:2012,Hartung:2012,Chang:2013,Shahinpour:2013,Wotzlaw:2014}.

A maximum clique algorithm can easily be adapted to find a maximum $k$-clique in a graph $G$ by
considering the graph $G^k$, which we describe below. However, it is not obvious that this is a
viable approach: even if $G$ is sparse, $G^k$ may not be, and the maximum clique problem on dense
graphs can be very challenging computationally. Here we take a state-of-the-art maximum clique
algorithm which is suitable for use on dense graphs \cite{Prosser:2012,McCreesh:2013}, and
investigate whether this approach is feasible in practice. We modify the algorithm to include a new
lazy ``global domination'' inference step---this technique provides no benefit for typical maximum
clique problems, but for maximum $k$-clique graphs it sometimes gives improvements of several orders
of magnitude. We present computational results for the maximum $k$-clique problem on a range of
benchmark and real-world graphs. We finish with a detailed look at random graphs.

Throughout, our graphs are finite, undirected, and contain no loops. If $G$ is a graph with vertex
set $V$ and edge set $E$, we may write $\vertexset(G)$ to mean $V$. The \emph{neighbourhood} of a
vertex $v$ in a graph $G$, written $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$.
The \emph{degree} of a vertex is the cardinality of its neighbourhood. The density of a graph,
denoted $D$, is the proportion of distinct pairs of vertices which have an edge between them. The
subgraph \emph{induced by} a set of vertices $W$ is the subgraph with vertex set $W$, and all edges
from the original graph that are between pairs of vertices in $W$. If $A$ and $B$ are sets of
vertices, we write $A \setminus B$ for the set of vertices which are in $A$ but not $B$, and we
write $A + v$ and $A - v$ for $A \cup \{v\}$ and $A \setminus \{v\}$ respectively.

\section{Algorithms}

Our approach for finding a maximum $k$-clique is presented as \cref{algorithm:maxKClique}. Our first
step (\mcline{power}) is to replace our input graph $G$ with a modified graph $G^k$. This graph has
the same vertex set as $G$, and edges between any two distinct vertices $v_1$ and $v_2$ iff there is
a path of length at most $k$ between $v_1$ and $v_2$ in $G$. We may construct this graph using a
bounded breadth-first search: we refer to Chang et al.\ \cite{Chang:2013} for how to implement this
quickly in practice. Now it is easy to see that maximum cliques in $G^k$ correspond with maximum
$k$-cliques in $G$ \cite{Balasundaram:2005}.

\begin{algorithm}\DontPrintSemicolon
    \begin{tikzpicture}[remember picture,overlay]
        \coordinate (cvrej1c) at ($(pic cs:cvrej1) + (0, 0.15)$);
        \coordinate (cvrej2c) at ($(pic cs:cvrej2) + (0, 0.00)$);
        \node [fill=safelightblue, rounded corners, fit=(cvrej1c) (cvrej2c)] { };

        \coordinate (cdom1c) at ($(pic cs:cdom1) + (0, 0.15)$);
        \coordinate (cdom2c) at ($(pic cs:cdom2) + (0, -0.10)$);
        \node [fill=safelightblue, rounded corners, fit=(cdom1c) (cdom2c)] { };

        \coordinate (cvinp1c) at ($(pic cs:cvinp1) + (0, 0.15)$);
        \coordinate (cvinp2c) at ($(pic cs:cvinp2) + (1, 0.00)$);
        \node [fill=safelightblue, rounded corners, fit=(cvinp1c) (cvinp2c)] { };

        \coordinate (cvrejg1c) at ($(pic cs:cvrejg1) + (0, 0.15)$);
        \coordinate (cvrejg2c) at ($(pic cs:cvrejg2) + (0, 0.00)$);
        \node [fill=safelightblue, rounded corners, fit=(cvrejg1c) (cvrejg2c)] { };

    \end{tikzpicture}

    \nl $\FuncSty{maximumKClique}$ :: (Graph $G$, Integer $k$) $\rightarrow$ Vertex Set \;
    \nl \Begin{
        \nl $G$ $\gets$ $G^k$ \mclabel{power} \;
        \nl permute $G$ so that vertices are in non-increasing degree order \mclabel{permute} \;
        \nl $\KwSty{global}$ $\Cbest$ $\gets$ $\emptyset$ \mclabel{cbest} \;
        \nl $\expand$($\emptyset$, $\vertexset(G)$) \mclabel{c} \mclabel{p} \;
        \nl $\KwSty{return}$ $\Cbest$ (unpermuted) \;
    }
    \;

    \nl $\expand$ :: (Vertex Set $C$, Vertex Set $P$) \;
    \nl \Begin{
        \nl ($\order$, $\bounds$) $\gets$ $\colourOrder$($P$) \mclabel{colouring} \;
        \nl \tikzmark{cvrej1}$\vrej$ $\gets$ $\kwunset$ \tikzmark{cvrej2} \;
        \nl \For{$i$ $\gets$ $|P|$ $\KwSty{downto}$ 1\mclabel{righttoleft}}{
            \nl \lIf{\textnormal{$|C|$ + $\bounds[i]$ $\le$ $|\Cbest|$}\mclabel{bound}}{$\KwSty{return}$}
            \nl \tikzmark{cdom1}\If{\textnormal{$\vrej$ $\ne$ $\kwunset$}}{
                \nl $P$ $\gets$ $P$ $\setminus$ $\{ w \in \vertexset(G) :
                \neighbourhood_G(w) - \vrej \subseteq \neighbourhood_G(\vrej) - w \}$
                \tikzmark{cdom2} \mclabel{dominated} \;
            }
            \nl $v$ $\gets$ $\order[i]$ \mclabel{v} \;
            \nl \tikzmark{cvinp1}\If{$v$ $\in$ $P$\tikzmark{cvinp2}\mclabel{alreadyrejected}}{
                \nl $C$ $\gets$ $C + v$ \mclabel{vinstart} \;
                \nl \lIf{\textnormal{$|C| > |\Cbest|$}}{$\Cbest$ $\gets$ $C$\mclabel{unseat}}
                \nl $P'$ $\gets$ $P$ $\cap$ $\neighbourhood_G(v)$ \mclabel{pprime} \;
                \nl \lIf{$P'$ $\ne$ $\emptyset$}{$\expand$($C$, $P'$)\mclabel{vinend}\mclabel{recurse}}
                \nl \tikzmark{creject1}$C$ $\gets$ $C - v$ \mclabel{vnotinstart} \;
                \nl $P$ $\gets$ $P - v$\tikzmark{creject2} \mclabel{vnotinend} \;
            }
            \nl \tikzmark{cvrejg1}$\vrej$ $\gets$ $v$ \tikzmark{cvrejg2} \mclabel{vrej} \;
        }
    }
    \;

    \nl $\colourOrder$ :: (Vertex Set $P$) $\rightarrow$ (Vertex Array, Int Array) \;
    \nl \Begin{
        \nl ($\order$, $\bounds$) $\gets$ ($[]$, $[]$) \;
        \nl $\uncoloured$ $\gets$ $P$ \;
        \nl $\colour$ $\gets$ $1$ \mclabel{colour1} \;
        \nl \While{$\uncoloured$ $\ne$ $\emptyset$\mclabel{uncoloured}}{
            \nl $\colourable$ $\gets$ $\uncoloured$ \mclabel{currentcolourstart} \;
            \nl \While{$\colourable$ $\ne$ $\emptyset$}{
                \nl $v$ $\gets$ the first vertex of $\colourable$ \;
                \nl append $v$ to $\order$ \;
                \nl append $\colour$ to $\bounds$ \;
                \nl $\uncoloured$ $\gets$ $\uncoloured - v$ \;
                \nl $\colourable$ $\gets$ $\colourable \setminus \neighbourhood_G(v)$ \mclabel{currentcolourend}\mclabel{notcolourable}\;
            }
            \nl $\colour$ $\gets$ $\colour + 1$ \mclabel{colourplus1}
        }
        \nl $\KwSty{return}$ ($\order$, $\bounds$) \;
    }

    \caption{An algorithm for the maximum $k$-clique problem.}
    \label{algorithm:maxKClique}
\end{algorithm}

\paragraph{Colouring}

The current state-of-the-art for the maximum clique problem on dense graphs, due to Tomita et al.\
\cite{Tomita:2003,Tomita:2007,Tomita:2010}, is to use branch and bound with a greedy graph
colouring.  A \emph{colouring} of a graph is an assignment of colours to vertices, such that
adjacent vertices are given different colours; if we can colour a graph using $c$ colours, then the
graph cannot contain a clique of size greater than $c$ (each vertex in a clique must be given a
different colour).

Obtaining a minimal colouring is \NP-hard, but we may create a greedy colouring in polynomial time.
This is done by the $\colourOrder$ routine: we start the first colour (\mcline{colour1}), and while
there are uncoloured vertices remaining (\mcline{uncoloured}), we try to give each vertex in turn
the current colour (\mclinerange{currentcolourstart}{currentcolourend}).  When we cannot colour any
further vertices, we start a new colour (\mcline{colourplus1}).

The key step in Tomita's algorithms is to produce a constructive colouring, which is used in a
clever way.  The $\colourOrder$ routine does not just return the number of colours used. Instead, it
returns a pair of arrays, $\order$ and $\bounds$. The $\order$ array contains vertices, in the order
in which they were coloured. The $i$th entry of the $\bounds$ array contains the colour number used
for the $i$th vertex in $\order$. We illustrate this in \cref{figure:colours}. Crucially, $\bounds$
is non-decreasing (i.e.\ $\bounds[i+1] \ge \bounds[i]$), and we may colour the subgraph induced by
the first $i$ vertices of $\order$ using $\bounds[i]$ colours.

\begin{figure}[b]
    \centering
    \begin{tikzpicture}
        % \input{figure-colour}
    \end{tikzpicture}

    \caption{The graph on the left has been coloured greedily, using four colours: vertices 1, 3, 4
        then 6 were given the first colour, then vertices 2 then 7 were given the second colour,
        then vertex 5 was given the third colour, and vertex 8 the fourth colour. The $\order$ array
        contains the vertices in the order they were coloured; the $i$th entry of the $\bounds$ array
        contains the number of colours used to colour the first $i$ vertices of $\order$.}
    \label{figure:colours}
\end{figure}

The order in which vertices are selected for colouring can have a large effect upon performance.
Various initial vertex orderings have been considered for the maximum clique problem---we refer to a
computational study by Prosser for details \cite{Prosser:2012}. Here we will colour vertices in a
static non-increasing degree order, which we do by permuting the graph at the top of search
(\mcline{permute}). We will \emph{not} be using Tomita et al.'s dynamic tie-breaking mechanism
\cite{Tomita:2010}: although doing so can sometimes be beneficial for small dense graphs in a
maximum clique context, for the larger graphs we will be considering here the cubic cost is
prohibitively expensive. For the same reason, we use a simple greedy colouring and do not use Tomita
et al.'s (cubic) colour repair step \cite{Tomita:2010}.

\paragraph{Branching and recursing}

We may now describe the main recursive part of the algorithm. If $v$ is a vertex, then a clique in
$G^k$ either contains only $v$ and vertices adjacent to $v$, or does not contain $v$. This allows us
to grow cliques by repeatedly picking a vertex, and branching upon whether or not to include it. Our
growing clique is stored in the variable $C$, which is initially empty (\mcline{c}). We also track
which vertices may still be added to $C$ in the variable $P$, which initially contains every vertex
(\mcline{p}). The $\expand$ procedure picks a vertex $v$ (\mcline{v}), then considers adding $v$ to
$C$ (\mclinerange{vinstart}{vinend}): we create a new $P'$ from $P$ (\mcline{pprime}) by
rejecting vertices which are not adjacent to $v$ (and thus every vertex in $P'$ is adjacent to
\emph{every} vertex in $C$). If vertices remain in $P'$, we recurse (\mcline{recurse}). We then take
the opposite branch choice, and consider rejecting from $P$ and $C$
(\mclinerange{vnotinstart}{vnotinend}). We then loop, and pick a new $v$.

\paragraph{Integrating the colour bound}

We keep track of the best solution we have found so far, which we call the \emph{incumbent}; this is
stored in $\Cbest$, which is initially empty (\mcline{cbest}). Whenever we find a new clique, we
compare its size to that of $\Cbest$, and if it is better, the incumbent is unseated
(\mcline{unseat}). Now we may make use of the colour bound. At the start of the recursive procedure
(\mcline{colouring}), we use $\colourOrder$ to produce a constructive greedy colouring of the
subgraph induced by $P$ into the array $\order$, with the colour numbers placed in $\bounds$. When
selecting $v$, we iterate over $\bounds$ from right to left (\mcline{righttoleft}). Now on
\mcline{bound} we know that the largest possible clique we could find at the current location has
size no greater than $|C| + \bounds[i]$, so if this cannot unseat the incumbent then we may abandon
search and backtrack.

\paragraph{Lazy global domination}

Aside from the $G^k$ step, what we have described so far is a standard maximum clique algorithm, and
all we have done is opted out of certain more computationally expensive inference steps (more
complicated initial vertex orderings, and cubic colourings). If we ignore the lines shown in blue,
we obtain the maximum clique algorithm variation that Prosser \cite{Prosser:2012} calls ``MCSa1''.
Now we will introduce a new lazy global domination rule which performs additional inference during
search. This rule is not specific to the maximum $k$-clique problem, and is also valid for the
maximum clique problem.

Let $v$ and $w$ be distinct vertices in a graph $G$ (they may or may not be adjacent). We say that
$v$ \emph{dominates} $w$ if the neighbourhood of $w$, excluding $v$, is a (possibly non-strict)
subset of the neighbourhood of $v$, excluding $w$. From a maximum clique perspective, this means
that $v$ is ``better than'' $w$. If $v$ and $w$ are adjacent, any clique containing $w$ may always
be extended by the inclusion of $v$; if $v$ and $w$ are non-adjacent, replacing $w$ with $v$ in any
clique containing $w$ cannot reduce the amount by which the clique may be grown.

Suppose a graph does contain one or more pairs of dominating vertices. We could make use of this
fact during search in at least two ways. Firstly, when accepting a vertex $w$, we may also
unconditionally accept any vertex $v$ which dominates $w$. Secondly, when rejecting a vertex $v$, we
may also unconditionally reject any vertex $w$ which is dominated by $v$.  We could also choose to
calculate domination globally (i.e.\ with respect to $G^k$, or even the original $G$), or locally
(i.e.\ with respect to the subgraph of $G^k$ induced by $C \cup P$).

Detecting whether one vertex dominates another may be done in linear time (we discuss this further
below), but finding all vertices dominated by a particular vertex is quadratic, and finding all
dominations is cubic. This is a heavy price to pay, if there are no dominating vertices.  This is
why such a rule has not previously been used in the maximum clique context: in the authors'
experience, most graphs typically considered for the maximum clique problem do not contain
dominating vertices, and those that do are too easy computationally for the step to be worthwhile.

However, some of the graphs we consider in the following section \emph{do} contain dominating
vertices, and although the maximum clique problem is trivial on these graphs, the maximum $k$-clique
problem is not for some values of $k$. Preliminary experiments suggested that the use of a
domination rule could be extremely beneficial in certain circumstances, but that in cases where it
had little effect, doing such a calculation introduced a substantial penalty to runtimes. Moreover,
even in graphs where dominating vertices are present, knowing this fact is sometimes not useful: it
is common for an optimal solution to be found straight away, and for the bound to be strong enough
to prove optimality immediately, so no branching occurs.

This motivates the design of a lazy global domination rule. We perform our domination checks
globally, with respect to $G^k$ (which may contain more dominating vertices than $G$), and we
remember and reuse the results of any domination checks we perform. We also only perform inference
on the ``reject'' case, to avoid introducing any cost when a solution is found and proven optimal
without branching.

The lines marked in blue in \cref{algorithm:maxKClique} show how this is done. When a vertex $\vrej$
is rejected, we remove from $P$ any vertex that is dominated (with respect to $G^k$) by $\vrej$.
This is \mcline{dominated}; the set of dominated vertices calculated here should be cached.  One
might expect that this calculation would appear after \mcline{vnotinend}. However, this introduces a
cost if the bound allows the next choice of $v$ to be eliminated. Thus we simply remember that we
have rejected $v$ by storing it in $\vrej$ (\mcline{vrej}), and lazily postpone the filtering until
after the bound has been checked.

Finally, note that we do not perform a new colouring when we reject dominated vertices---doing so
typically does not lead to a smaller bound, since most colour classes contain many vertices. Thus
when we select a $v$ from $\order$, it is now possible that $v$ has already been rejected. We check
for this on line \mcline{alreadyrejected}.

\paragraph{Bitset encoding} San Segundo et al.\ \cite{SanSegundo:2011,SanSegundo:2013} observed that
the performance of Tomita's algorithms could be enhanced substantially by using a bitset encoding to
obtain a form of SIMD-like parallelism, without altering the steps taken. We have taken such an
approach here too, although we do not describe it explicitly---when permuting $G$ on line
\mcline{permute}, the graph should be re-encoded as an array of adjacency bitsets. (It is not
helpful to do this before constructing $G^k$.) Now the intersection on \mcline{pprime} becomes a
simple bitwise ``and'' operation, and the intersection with complement on \mcline{notcolourable} is
a bitwise ``and not'' operation. This is beneficial when testing for dominance, too: each bit in the
dominated set on \mcline{dominated} may be determined by a bitwise ``and not'', unsetting a bit, and
testing whether the result is empty; the set difference is again a bitwise ``and not'' operation.

\section{Experimental Results}

Here we give experimental results on a range of standard benchmarks, and on real-world and random
graphs. Where timing results are reported, the experiments were run on a machine with Intel E5645
processors, and single-threaded runtimes are given. The time taken to read in the graph from a file
is excluded, but preprocessing time (including the construction of $G^k$ and the bitset encoding) is
included. We use the term \emph{nodes} to refer to the number of recursive calls made by the
branch-and-bound part of the algorithm.

\subsection{Real-World Graphs}

We begin with a selection of real-world and standard benchmark graphs. We look at $k$ equal to 2, 3
and 4 in every case---this is a standard practice for the $k$-club problem
\cite{Chang:2013,Wotzlaw:2014}.

\afterpage{
    {
    \onecolumn
\setlength{\LTcapwidth}{\textwidth}
\centering
\begin{longtable}{l c r rrr rrr}
    \caption{Experimental results for a range of graphs. For each graph, we consider $k$ equal to 2,
        3 and 4. In each case we show the density of $G^k$, and then for both the unmodified
        algorithm and the algorithm with our lazy global domination step, we give the size of a
        maximum $k$-clique, the number of nodes required, and the runtime in seconds. Some results
        were aborted after one day.}
    \label{table:results}
    \\

    \toprule

    & & & \multicolumn{3}{c}{Unmodified} & \multicolumn{3}{c}{With Domination} \\
    \cmidrule(lr){4-6}
    \cmidrule(lr){7-9}
    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} &
    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
    \midrule

    \endfirsthead

    \caption[]{(continued from previous page)} \\
    \toprule

    & & & \multicolumn{3}{c}{Unmodified} & \multicolumn{3}{c}{With Domination} \\
    \cmidrule(lr){4-6}
    \cmidrule(lr){7-9}
    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} &
    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
    \midrule

    \endhead

    \midrule
    \endfoot

    \bottomrule
    \endlastfoot

    % \inputhaxx{gen-table-erdos}
    \midrule
    % \inputhaxx{gen-table-dimacs}
    \midrule
    % \inputhaxx{gen-table-dimacs10walshaw}
    \midrule
    % \inputhaxx{gen-table-dimacs10cluster}
\end{longtable}}\twocolumn}

\paragraph{Erd\H{o}s collaboration graphs}

In the first part of \cref{table:results} we present experimental results from Erd\H{o}s
collaboration graphs from the Pajek dataset by Vladimir Batagelj and Andrej
Mrvar\footnote{\texttt{http://vlado.fmf.uni-lj.si/pub/networks/data/}}.  We were able to solve all of
these problems in under eight minutes (and all but three in under two seconds) when using the
domination rule. However, using the unmodified maximum clique algorithm, two of these results did
not finish running within one day. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.

In several cases, the algorithm found and proved an optimal solution immediately ($\tilde{\omega}_k$
is equal to the number of search nodes). This illustrates the necessity of laziness: if we simply
computed dominating pairs upfront, we would be paying a cubic preprocessing cost for an algorithm
which is effectively quadratic in practice.

By comparing these results with the $k$-club results of Chang et al.\ \cite{Chang:2013}, we see that
in all but four cases the $k$-clique and $k$-club numbers are equal; all of these differences occur
when $k = 4$. (Chang et al.\ did not investigate the ``Erdos02'' graph, but Wotzlaw
\cite{Wotzlaw:2014} confirmed privately that the $k$-clique and $k$-club numbers are the same here
too.) On the other hand, the $k$-clique numbers are sometimes much easier to find, both
algorithmically and computationally.

\paragraph{Clique graphs}

In the second part of \cref{table:results} we present results from the ``clique'' graphs from the
Second DIMACS implementation challenge\footnote{\texttt{http://dimacs.rutgers.edu/Challenges/}}. These
graphs were designed to test maximum clique implementations. Nearly all of these graphs have
diameter 2, so a 2-clique covers the entire graph---we have ignored these. The only exceptions are
the ``c-fat'' family (all of which are trivial for a maximum clique solver), and one of the
``p\_hat'' graphs.

With the domination rule, we solve all of these problems within a tenth of a second. Without, two of
the results take over a day, and the rest remain trivial. Note that in several cases, for some
values of $k$ a $k$-clique covers the entire graph.  Again using Chang et al.'s results
\cite{Chang:2013}, we see that for the first six graphs in this table the $k$-clique and $k$-club
numbers are the same for each value of $k$ (Chang et al.\ did not investigate ``c-fat500-10'' or
``p-hat300-1'').

\paragraph{Partitioning graphs}

The third part of \cref{table:results} presents results from the smallest 20 partitioning graphs
from the 10th DIMACS Implementation
Challenge\footnote{\texttt{http://staffweb.cms.gre.ac.uk/~wc06/partition/}}.  Many of these graphs are
considerably larger than those typically considered for the maximum clique problem, and we might
expect our $O(|V|^2)$ memory requirements to cause problems. Nonetheless, with the domination rule
there is only one instance which we were unable to solve within a day (and without the domination
rule, there are two).

On the other hand, we sometimes see a significant cost where the domination rule does not help, and
where the proof of optimality is not immediate: in ``3elt'' and ``4elt'', our runtimes can nearly
double, and for ``cs4'' and ``cti'' the slowdown is sometimes over a factor of ten. In other words,
laziness does not help when the rule turns out to be used, but useless.

Five of these graphs were considered for the $k$-club problem by Wotzlaw \cite{Wotzlaw:2014}. In all
five cases, the $k$-clique and $k$-club numbers are the same for $k$ equal to 2, 3 and 4. However,
the $k$-clique number was again consistently much easier to find.

\paragraph{Clustering graphs}

The final part of \cref{table:results} presents results from the smallest 20 partitioning graphs
from the 10th DIMACS Implementation
Challenge\footnote{\texttt{http://www.cc.gatech.edu/dimacs10/archive/clustering.shtml}}. Again, from a
maximum clique perspective these would be considered unusually large graphs. However, only five were
unsolvable within a day (plus a sixth when the domination rule was not used), and half of the
problems took under two seconds.

Seven of these graphs were considered for the $k$-club problem by Wotzlaw \cite{Wotzlaw:2014}. In
these cases, the $2$-clique and $2$-club numbers are the same, except for ``football'' where the
$2$-club number is 16 but the $2$-clique number is 17; for $k = 3$ and $k = 4$ there are some
differences. The difference in computational difficulty between the $k$-clique and $k$-club problems
really stands out here: for ``polblogs'' with $k = 3$ and $k = 4$, Wotzlaw was unable to prove
optimality within an hour, but we required less than a second to do so. In both of these cases the
$k$-clique and $k$-club numbers are the same, which suggests a potential improvement to $k$-club
algorithms: first solve the maximum $k$-clique problem instead, and test whether the result found is
a $k$-club, before embarking upon a more complicated search. (Note that a negative result does not
imply that the $k$-clique and $k$-club numbers necessarily differ, since solutions are not unique.)

\subsection{Random Graphs}

An Erd\H{o}s-R\'{e}nyi random graph $G(n, p)$ has $n$ vertices, and an edge between each distinct
pair of vertices with probability $p$, chosen independently. Here we investigate the size of a
maximum $k$-clique in such graphs, and the complexity of finding it. In each case, we use an average
over 100 samples for every point.  We do not use the domination rule for these experiments: the
probability of random graphs having dominating vertices is very low.

In \cref{figure:graph-omega} we illustrate the average value of $\tilde{\omega}_k$ in $G(200,
p)$ for different values of $k$, and a range of values of $p$ for the $x$-axis. We see that even for
very low edge probabilities, a maximum $k$-clique quickly covers the entire graph.  (This is in contrast to
the maximum clique problem, where a maximum clique does not even cover a quarter of the graph for
edge probabilities below 0.75.) In \cref{figure:graph-nodes} we show the average size of the
search space (number of nodes, or recursive calls made) for the same problem. We see that there is a
complexity peak for each $k$, although the peak is much smaller for $k = 4$ than it is for $k = 3$,
which is in turn much smaller than it is for $k = 2$. The peak also occurs for lower edge
probabilities as $k$ increases. For contrast, for the maximum clique problem, the peak occurs at
around edge probability 0.9, and is two orders of magnitude larger.

\begin{figure}[p] % these two on the same page please
    \centering
    % \input{gen-graph-omega}
    \caption{Values of $\tilde{\omega}_k$ for random graphs $G(200, p)$, with varying edge probabilities.
        We see that even for very low edge probabilities, a
    maximum $k$-clique quickly covers the entire graph. This is in contrast to maximum cliques, which remain
    small even at much higher edge probabilities.}
    \label{figure:graph-omega}

    \vspace{1.5em}

    % \input{gen-graph-nodes}
    \caption{Search space size for random graphs $G(200, p)^k$, with varying edge probabilities. We see that
        $4$-clique is easier than $3$-clique in practice, which in turn is easier than $2$-clique. (The
        complexity peak for maximum clique occurs at around edge probability 0.9, and requires
        approximately 15 million search nodes.)}
    \label{figure:graph-nodes}
\end{figure}

In \cref{figure:graph-omega-2,figure:graph-nodes-2} we show the effect of
changing $n$ and fixing $k = 2$. As $n$ increases from 50 to 200, the complexity peak becomes much
more pronounced, and shifts slightly towards the left (lower edge probabilities).

\begin{figure}[p] % these two on the same page please
    \centering
    % \input{gen-graph-omega-2}
    \caption{The size of a maximum $2$-clique in random graphs $G(n, p)$ with varying edge
        probabilities, and different values of $n$. For $G(50, p)$, a $2$-clique has size average 50
    from $p = 0.42$ onwards.}
    \label{figure:graph-omega-2}

    \vspace{1.5em}

    % \input{gen-graph-nodes-2}
    \caption{The search space size for the maximum $2$-clique problem in random graphs $G(n, p)$
        with varying edge probabilities, and for different values of $n$. As $n$ increases, the
    complexity peak grows and moves slowly to the left.}
    \label{figure:graph-nodes-2}
\end{figure}

\section{Conclusion}

We have shown that using a maximum clique algorithm to solve the maximum $k$-clique algorithm for a
graph $G$ by considering $G^k$ in place of $G$ is feasible in practice. This is despite $G^k$
potentially being dense even if $G$ is sparse.

We introduced a new lazy global domination rule. This was sometimes extremely beneficial---without
this rule, we would have been unable to solve six of the problem instances we considered, and many
others would have taken much longer. However, even with laziness there is still sometimes a cost to
pay when this rule does nothing. This rule is thus harmful for the graphs typically considered for
the maximum clique problem, and we see the benefit of tailoring algorithms to the problem being
solved. We suggest that a similar rule may also be useful for the maximum $k$-club problem.

Quite often, we saw $k$-clique numbers and $k$-club numbers being the same. However, solving the
maximum $k$-clique problem is much easier, both in terms of the algorithm and computationally. Thus
it is worth checking whether the simpler model would be sufficient for practical applications before
trying to solve the $k$-club problem.

In random graphs, we saw that $G(n, p)^k$ is easier than $G(n, p')$ with some higher probability
$p'$. We also saw that as $k$ increases, the problem gets easier---this was not typically the case
for some of the real world graphs.

Our results suggest that $k$ is a very coarse grained parameter. We saw that often a $2$-clique or
$3$-clique would cover the entire graph. In these circumstances the increased restrictions for
$k$-club are of no benefit. It is not obvious if somehow allowing a ``fractional'' value of $k$
could give more fine-grained control. Thus it may be worth considering other clique relaxations not
based upon distance (although other models also have problems: a density-based relaxation known as
quasi-clique, for example, can allow vertices with only a single edge to be added to a ``clique''
\cite{Abello:2002}).

\bibliographystyle{aaai}
\bibliography{paper}

\end{document}

### 2016-03-29 paper/paper.tex
Bring in the pretty things

@@
 \usepackage{afterpage}
+\usepackage{siunitx}
+\usepackage{tikz}                      % For pretty pictures
+\usepackage{gnuplot-lua-tikz}          % graphs
+
+% \usepackage{showframe}
 
@@
 decorations.pathreplacing, patterns, tikzmark, fit}
+
+% \input isn't expandable. yay.
+\makeatletter\newcommand*\inputhaxx[1]{\@@input #1}\makeatother
 
@@
 
-\frenchspacing
-\setlength{\pdfpagewidth}{8.5in}
-\setlength{\pdfpageheight}{11in}
-\pdfinfo{
-/Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
-/Author (Ciaran McCreesh, Patrick Prosser)}
-\setcounter{secnumdepth}{0}
+\crefname{algocf}{algorithm}{algorithm}
+\Crefname{algocf}{Algorithm}{Algorithms}
 
@@
 \newcommand{\order}{\mathit{order}}
-\newcommand{\colour}{\mathit{colour}}
+\newcommand{\colour}{\mathit{currentColour}}
 \newcommand{\uncoloured}{\mathit{uncoloured}}
@@
 \newcommand{\expand}{\FuncSty{expand}}
-\newcommand{\colourOrder}{\FuncSty{colourOrder}}
+\newcommand{\dominated}{\FuncSty{dominated}}
+\newcommand{\colourOrder}{\FuncSty{colour}}
 \newcommand{\vrej}{v_{\mathit{rej}}}
@@
 
+% drawing styles
+\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.4cm, font=\small}}
+\tikzset{notvertex/.style={vertex, color=white, text=black}}
+\tikzset{plainvertex/.style={vertex}}
+\tikzset{selectedvertex/.style={vertex, fill=safedarkblue}}
+\tikzset{delvertex/.style={vertex, dotted, color=safelightgrey}}
+\tikzset{vertexc1/.style={vertex, fill=safelightblue}}
+\tikzset{vertexc2/.style={vertex, fill=safedarkblue, text=safenearlywhite}}
+\tikzset{vertexc3/.style={vertex, fill=safelightorange}}
+\tikzset{vertexc4/.style={vertex, fill=safedarkorange, text=safenearlywhite}}
+
+\tikzset{edge/.style={color=safelightgrey}}
+\tikzset{bedge/.style={very thick}}
+\tikzset{deledge/.style={dotted, color=safelightgrey}}
+\tikzset{edgel1/.style={color=safeverydarkblue}}
+\tikzset{edgel2/.style={color=safemediumorange}}
+\tikzset{edgel3/.style={very thick, color=safedarkblue}}
+\tikzset{edgel4/.style={very thick, color=safelightorange}}
+
+\tikzset{processarrow/.style={->, very thick, decorate, decoration={snake, post length=0.5mm}}}
+\tikzset{brace/.style={decorate, decoration={brace}, very thick}}
+\tikzset{label/.style={font=\small}}
+
+\makeatletter\@ifpackageloaded{showframe}{
+    % show lines down the middle columns too
+    % http://tex.stackexchange.com/questions/16199/test-if-a-package-or-package-option-is-loaded
+\newlength\Fcolumnseprule\setlength\Fcolumnseprule{0.4pt}\def\@outputdblcol{%
+  \if@firstcolumn\global \@firstcolumnfalse \global \setbox\@leftcolumn \box\@outputbox
+  \else
+    \global \@firstcolumntrue
+    \setbox\@outputbox \vbox{\hb@xt@\textwidth{\hb@xt@\columnwidth{\box\@leftcolumn \hss}%
+    \vrule \@width\Fcolumnseprule\hfil{\normalcolor\vrule \@width\columnseprule}
+    \hfil\vrule \@width\Fcolumnseprule\hb@xt@\columnwidth {\box\@outputbox \hss}}}\@combinedblfloats\@outputpage
+    \begingroup\@dblfloatplacement\@startdblcolumn\@whilesw\if@fcolmade \fi{\@outputpage\@startdblcolumn}\endgroup
+  \fi
+}}{}
+\makeatother
+
+% \frenchspacing
+\setlength{\pdfpagewidth}{8.5in}
+\setlength{\pdfpageheight}{11in}
+\pdfinfo{
+/Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
+/Author (Ciaran McCreesh, Patrick Prosser)}
+\setcounter{secnumdepth}{0}
+
 \begin{document}
-% The file aaai.sty is the style file for AAAI Press
-% proceedings, working notes, and technical reports.
-%
 \title{Finding Maximum $k$-Cliques Faster using Lazy Global Domination}
@@
 \maketitle
+
 \begin{abstract}
@@
 \section{Introduction}
-
-\begin{figure}[b] % first page please
-    \centering
-    \begin{tikzpicture}
-        % \input{figure-k-clique-1}
-    \end{tikzpicture}\hspace{4em}\begin{tikzpicture}
-        % \input{figure-k-clique-2}
-    \end{tikzpicture}
-
-    \caption{On the left, a graph, with its unique maximum clique $\{1, 2, 5, 8\}$ of size 4
-        highlighted. On the right, the same graph, with a maximum $2$-clique $\{1, 2, 3, 4, 5, 6, 8\}$
-        of size 7 highlighted. This is not a $2$-club, since the only path of length 2 between vertices 3 and 6
-        goes through vertex 7. A $3$-clique covers the entire graph.}
-    \label{figure:k-cliques}
-\end{figure}
 
@@
 is a $k$-clique).
+
+\begin{figure}[b!] % first page please
+    \centering
+    \begin{tikzpicture}[scale=0.8]
+        \newcount \c
+        \foreach \n in {1, ..., 8}{
+            \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+            \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 5 \OR \n = 8}{
+                \node[selectedvertex] (N\n) at (\the\c:1.2) {\n};
+            }{
+                \node[plainvertex] (N\n) at (\the\c:1.2) {\n};
+            }
+        }
+
+        \draw [edge] (N2) -- (N3);
+        \draw [edge] (N2) -- (N4);
+        \draw [edge] (N3) -- (N7);
+        \draw [edge] (N4) -- (N5);
+        \draw [edge] (N5) -- (N6);
+        \draw [edge] (N6) -- (N7);
+
+        \draw [bedge] (N1) -- (N2);
+        \draw [bedge] (N1) -- (N5);
+        \draw [bedge] (N1) -- (N8);
+        \draw [bedge] (N2) -- (N5);
+        \draw [bedge] (N2) -- (N8);
+        \draw [bedge] (N5) -- (N8);
+    \end{tikzpicture}\hspace{3em}\begin{tikzpicture}[scale=0.8]
+        \newcount \c
+        \foreach \n in {1, ..., 8}{
+            \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+            \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 3 \OR \n = 4 \OR \n = 5 \OR \n = 6 \OR \n = 8}{
+                \node[selectedvertex] (N\n) at (\the\c:1.2) {\n};
+            }{
+                \node[plainvertex] (N\n) at (\the\c:1.2) {\n};
+            }
+        }
+
+        \draw [edge] (N3) -- (N7);
+        \draw [edge] (N6) -- (N7);
+
+        \draw [bedge] (N1) -- (N2);
+        \draw [bedge] (N1) -- (N5);
+        \draw [bedge] (N1) -- (N8);
+        \draw [bedge] (N2) -- (N5);
+        \draw [bedge] (N2) -- (N8);
+        \draw [bedge] (N5) -- (N8);
+        \draw [bedge] (N2) -- (N3);
+        \draw [bedge] (N2) -- (N4);
+        \draw [bedge] (N4) -- (N5);
+        \draw [bedge] (N5) -- (N6);
+    \end{tikzpicture}
+
+    \caption{On the left, a graph, with its unique maximum clique $\{1, 2, 5, 8\}$ of size 4
+        highlighted. On the right, the same graph, with a maximum $2$-clique $\{1, 2, 3, 4, 5, 6, 8\}$
+        of size 7 highlighted. This is not a $2$-club, since the only path of length 2 between vertices 3 and 6
+        goes through vertex 7. A $3$-clique covers the entire graph.}
+    \label{figure:k-cliques}
+\end{figure}
 
@@
 
-    \nl $\FuncSty{maximumKClique}$ :: (Graph $G$, Integer $k$) $\rightarrow$ Vertex Set \;
+    \nl $\FuncSty{maxKClique}$ :: (Graph $G$, Integer $k$) $\rightarrow$ Vertex Set \;
     \nl \Begin{
@@
             \nl \tikzmark{cdom1}\If{\textnormal{$\vrej$ $\ne$ $\kwunset$}}{
-                \nl $P$ $\gets$ $P$ $\setminus$ $\{ w \in \vertexset(G) :
-                \neighbourhood_G(w) - \vrej \subseteq \neighbourhood_G(\vrej) - w \}$
-                \tikzmark{cdom2} \mclabel{dominated} \;
+                \nl $P$ $\gets$ $P$ $\setminus$ $\dominated(\vrej)$ \tikzmark{cdom2} \mclabel{dominated} \;
             }
@@
     }
+    \;
 
-    \caption{An algorithm for the maximum $k$-clique problem.}
+    \nl $\dominated$ :: (Vertex $v$) $\rightarrow$ Vertex Set \;
+    \nl \Begin{
+        \nl $\KwSty{return}$ $\{ w \in \vertexset(G) : \neighbourhood_G(w) - v \subseteq \neighbourhood_G(v) - w \} $\;
+    }
+    \caption{Solving the maximum $k$-clique problem.}
     \label{algorithm:maxKClique}
@@
     \centering
-    \begin{tikzpicture}
-        % \input{figure-colour}
+    \begin{tikzpicture}[scale=0.8]
+\newcount \c
+\foreach \n in {1, ..., 8}{
+    \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+    \ifthenelse{\n = 1 \OR \n = 3 \OR \n = 4 \OR \n = 6}{
+        \node[vertexc1] (N\n) at (\the\c:1.2) {\n};
+    }{}
+    \ifthenelse{\n = 2 \OR \n = 7}{
+        \node[vertexc2] (N\n) at (\the\c:1.2) {\n};
+    }{}
+    \ifthenelse{\n = 5}{
+        \node[vertexc3] (N\n) at (\the\c:1.2) {\n};
+    }{}
+    \ifthenelse{\n = 8}{
+        \node[vertexc4] (N\n) at (\the\c:1.2) {\n};
+    }{}
+}
+
+\draw [edge] (N2) -- (N3);
+\draw [edge] (N2) -- (N4);
+\draw [edge] (N3) -- (N7);
+\draw [edge] (N4) -- (N5);
+\draw [edge] (N5) -- (N6);
+\draw [edge] (N6) -- (N7);
+\draw [edge] (N1) -- (N2);
+\draw [edge] (N1) -- (N5);
+\draw [edge] (N1) -- (N8);
+\draw [edge] (N2) -- (N5);
+\draw [edge] (N2) -- (N8);
+\draw [edge] (N5) -- (N8);
+
+\draw [processarrow] (1.6, 0) -> (2.7, 0);
+
+\coordinate (Ms) at (3.0, 0.35);
+\node[right = 0.0 of Ms, vertexc1] (M1) {1};
+\node[right = 0.2 of M1, vertexc1] (M2) {3};
+\node[right = 0.2 of M2, vertexc1] (M3) {4};
+\node[right = 0.2 of M3, vertexc1] (M4) {6};
+\node[right = 0.2 of M4, vertexc2] (M5) {2};
+\node[right = 0.2 of M5, vertexc2] (M6) {7};
+\node[right = 0.2 of M6, vertexc3] (M7) {5};
+\node[right = 0.2 of M7, vertexc4] (M8) {8};
+
+\draw[brace] ($(M1.north west)+(0.0,0.3)$) -- ($(M8.north east)+(0.0,0.3)$);
+\node[anchor=south, label] at ($(M1)!0.5!(M8)$)[yshift=0.7cm] { Vertices in colour order };
+
+\coordinate (Bs) at (3.0, -0.35);
+\node[right = 0.0 of Bs, notvertex] (B1) {1};
+\node[right = 0.2 of B1, notvertex] (B2) {1};
+\node[right = 0.2 of B2, notvertex] (B3) {1};
+\node[right = 0.2 of B3, notvertex] (B4) {1};
+\node[right = 0.2 of B4, notvertex] (B5) {2};
+\node[right = 0.2 of B5, notvertex] (B6) {2};
+\node[right = 0.2 of B6, notvertex] (B7) {3};
+\node[right = 0.2 of B7, notvertex] (B8) {4};
+
+\draw[brace] ($(B8.south east)+(0.0,-0.2)$) -- ($(B1.south west)+(0.0,-0.2)$);
+\node[anchor=north, label] at ($(B1)!0.5!(B8)$)[yshift=-0.6cm] { Number of colours used };
     \end{tikzpicture}
 
-    \caption{The graph on the left has been coloured greedily, using four colours: vertices 1, 3, 4
-        then 6 were given the first colour, then vertices 2 then 7 were given the second colour,
-        then vertex 5 was given the third colour, and vertex 8 the fourth colour. The $\order$ array
-        contains the vertices in the order they were coloured; the $i$th entry of the $\bounds$ array
-        contains the number of colours used to colour the first $i$ vertices of $\order$.}
+    \caption{The graph on the left has been coloured greedily, using four
+        colours: vertices 1, 3, 4 then 6 were given the first colour, then
+        vertices 2 then 7 were given the second colour, then vertex 5 was given
+        the third colour, and vertex 8 the fourth colour. The $\order$ array,
+        on top, contains the vertices in the order they were coloured; the
+        $i$th entry of the $\bounds$ array, below, contains the number of colours
+        used to colour the first $i$ vertices of $\order$.}
     \label{figure:colours}
@@
 
-    % \inputhaxx{gen-table-erdos}
+    \inputhaxx{gen-table-erdos}
     \midrule
-    % \inputhaxx{gen-table-dimacs}
+    \inputhaxx{gen-table-dimacs}
     \midrule
-    % \inputhaxx{gen-table-dimacs10walshaw}
+    \inputhaxx{gen-table-dimacs10walshaw}
     \midrule
-    % \inputhaxx{gen-table-dimacs10cluster}
+    \inputhaxx{gen-table-dimacs10cluster}
 \end{longtable}}\twocolumn}
@@
 
-\begin{figure}[p] % these two on the same page please
+\begin{figure}[t]
     \centering
-    % \input{gen-graph-omega}
+    \input{gen-graph-omega}
     \caption{Values of $\tilde{\omega}_k$ for random graphs $G(200, p)$, with varying edge probabilities.
-        We see that even for very low edge probabilities, a
-    maximum $k$-clique quickly covers the entire graph. This is in contrast to maximum cliques, which remain
-    small even at much higher edge probabilities.}
+        We see that even for very low edge probabilities, a maximum $k$-clique
+    quickly covers the entire graph when $k > 1$.}
     \label{figure:graph-omega}
+\end{figure}
 
-    \vspace{1.5em}
-
-    % \input{gen-graph-nodes}
+\begin{figure}[t]
+    \centering
+    \input{gen-graph-nodes}
     \caption{Search space size for random graphs $G(200, p)^k$, with varying edge probabilities. We see that
@@
 
-\begin{figure}[p] % these two on the same page please
+\begin{figure}[t]
     \centering
-    % \input{gen-graph-omega-2}
+    \input{gen-graph-omega-2}
     \caption{The size of a maximum $2$-clique in random graphs $G(n, p)$ with varying edge
@@
     \label{figure:graph-omega-2}
+\end{figure}
 
-    \vspace{1.5em}
-
-    % \input{gen-graph-nodes-2}
+\begin{figure}[t]
+    \centering
+    \input{gen-graph-nodes-2}
     \caption{The search space size for the maximum $2$-clique problem in random graphs $G(n, p)$

### 2016-03-29 paper/paper.tex
Some more

@@
+% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :
+
 \def\year{2016}\relax
-%File: formatting-instruction.tex
 \documentclass[letterpaper]{article}
@@
 \usepackage[ruled,vlined]{algorithm2e} % algorithms
-\usepackage{longtable}
 \usepackage{booktabs}                  % less ugly tables
@@
 \usepackage{afterpage}
-\usepackage{siunitx}
+\usepackage[tight-spacing=true]{siunitx}
 \usepackage{tikz}                      % For pretty pictures
@@
 
+% cref style
 \crefname{algocf}{algorithm}{algorithm}
 \Crefname{algocf}{Algorithm}{Algorithms}
+\crefname{figure}{figure}{figures}
+\Crefname{figure}{Figure}{Figures}
+\crefname{table}{table}{tables}
+\Crefname{table}{Table}{Tables}
 
@@
 \makeatother
+
+\renewcommand{\topfraction}{0.9}
+\renewcommand{\bottomfraction}{0.8}
+\renewcommand{\dbltopfraction}{0.9}
+\renewcommand{\textfraction}{0.05}
+\setcounter{dbltopnumber}{2}
+\renewcommand{\floatpagefraction}{0.85}
+\renewcommand{\dblfloatpagefraction}{0.85}
 
@@
 
-\begin{algorithm}\DontPrintSemicolon
+\begin{algorithm}[t]\DontPrintSemicolon
     \begin{tikzpicture}[remember picture,overlay]
@@
         \nl $G$ $\gets$ $G^k$ \mclabel{power} \;
-        \nl permute $G$ so that vertices are in non-increasing degree order \mclabel{permute} \;
+        \nl permute $G$ into non-increasing degree order \mclabel{permute} \;
         \nl $\KwSty{global}$ $\Cbest$ $\gets$ $\emptyset$ \mclabel{cbest} \;
@@
     }
-    \;
+    \vspace{0.5ex}
 
@@
             \nl \tikzmark{cvinp1}\If{$v$ $\in$ $P$\tikzmark{cvinp2}\mclabel{alreadyrejected}}{
-                \nl $C$ $\gets$ $C + v$ \mclabel{vinstart} \;
-                \nl \lIf{\textnormal{$|C| > |\Cbest|$}}{$\Cbest$ $\gets$ $C$\mclabel{unseat}}
+                \nl $C'$ $\gets$ $C + v$ \mclabel{vinstart} \;
+                \nl \lIf{\textnormal{$|C'| > |\Cbest|$}}{$\Cbest$ $\gets$ $C'$\mclabel{unseat}}
                 \nl $P'$ $\gets$ $P$ $\cap$ $\neighbourhood_G(v)$ \mclabel{pprime} \;
-                \nl \lIf{$P'$ $\ne$ $\emptyset$}{$\expand$($C$, $P'$)\mclabel{vinend}\mclabel{recurse}}
-                \nl \tikzmark{creject1}$C$ $\gets$ $C - v$ \mclabel{vnotinstart} \;
-                \nl $P$ $\gets$ $P - v$\tikzmark{creject2} \mclabel{vnotinend} \;
+                \nl \lIf{$P'$ $\ne$ $\emptyset$}{$\expand$($C'$, $P'$)\mclabel{vinend}\mclabel{recurse}}
+                \nl $P$ $\gets$ $P - v$\tikzmark{creject2} \mclabel{vnotin} \;
             }
@@
 the opposite branch choice, and consider rejecting from $P$ and $C$
-(\mclinerange{vnotinstart}{vnotinend}). We then loop, and pick a new $v$.
+(\mcline{vnotin}). We then loop, and pick a new $v$.
 
@@
 fact during search in at least two ways. Firstly, when accepting a vertex $w$, we may also
-unconditionally accept any vertex $v$ which dominates $w$. Secondly, when rejecting a vertex $v$, we
-may also unconditionally reject any vertex $w$ which is dominated by $v$.  We could also choose to
-calculate domination globally (i.e.\ with respect to $G^k$, or even the original $G$), or locally
-(i.e.\ with respect to the subgraph of $G^k$ induced by $C \cup P$).
+unconditionally accept any vertex $v$ which both dominates and is adjacent to $w$. Secondly, when
+rejecting a vertex $v$, we may also unconditionally reject any vertex $w$ which is dominated by $v$.
+We could also choose to calculate domination globally (i.e.\ with respect to $G^k$, or even the
+original $G$), or locally (i.e.\ with respect to the subgraph of $G^k$ induced by $C \cup P$).
 
@@
 This is \mcline{dominated}; the set of dominated vertices calculated here should be cached.  One
-might expect that this calculation would appear after \mcline{vnotinend}. However, this introduces a
+might expect that this calculation would appear after \mcline{vnotin}. However, this introduces a
 cost if the bound allows the next choice of $v$ to be eliminated. Thus we simply remember that we
@@
 
-\afterpage{
-    {
-    \onecolumn
-\setlength{\LTcapwidth}{\textwidth}
-\centering
-\begin{longtable}{l c r rrr rrr}
-    \caption{Experimental results for a range of graphs. For each graph, we consider $k$ equal to 2,
-        3 and 4. In each case we show the density of $G^k$, and then for both the unmodified
-        algorithm and the algorithm with our lazy global domination step, we give the size of a
-        maximum $k$-clique, the number of nodes required, and the runtime in seconds. Some results
-        were aborted after one day.}
-    \label{table:results}
-    \\
-
-    \toprule
-
-    & & & \multicolumn{3}{c}{Unmodified} & \multicolumn{3}{c}{With Domination} \\
-    \cmidrule(lr){4-6}
-    \cmidrule(lr){7-9}
-    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} &
-    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
-    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
-    \midrule
-
-    \endfirsthead
-
-    \caption[]{(continued from previous page)} \\
-    \toprule
-
-    & & & \multicolumn{3}{c}{Unmodified} & \multicolumn{3}{c}{With Domination} \\
-    \cmidrule(lr){4-6}
-    \cmidrule(lr){7-9}
-    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} &
-    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
-    \multicolumn{1}{c}{$\tilde{\omega}_k$} & \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
-    \midrule
-
-    \endhead
-
-    \midrule
-    \endfoot
-
-    \bottomrule
-    \endlastfoot
-
-    \inputhaxx{gen-table-erdos}
-    \midrule
-    \inputhaxx{gen-table-dimacs}
-    \midrule
-    \inputhaxx{gen-table-dimacs10walshaw}
-    \midrule
-    \inputhaxx{gen-table-dimacs10cluster}
-\end{longtable}}\twocolumn}
-
 \paragraph{Erd\H{o}s collaboration graphs}
 
-In the first part of \cref{table:results} we present experimental results from Erd\H{o}s
+In \cref{table:erdos} we present experimental results from Erd\H{o}s
 collaboration graphs from the Pajek dataset by Vladimir Batagelj and Andrej
-Mrvar\footnote{\texttt{http://vlado.fmf.uni-lj.si/pub/networks/data/}}.  We were able to solve all of
+Mrvar\footnote{http://vlado.fmf.uni-lj.si/pub/networks/data/}.  We were able to solve all of
 these problems in under eight minutes (and all but three in under two seconds) when using the
@@
 not finish running within one day. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.
+
+\begin{table}
+    \tiny
+    \begin{tabular}{l c rr rr rr}
+        \toprule
+        & & & & \multicolumn{2}{c}{Unmodified} & \multicolumn{2}{c}{With Domination} \\
+    \cmidrule(lr){5-6}
+    \cmidrule(lr){7-8}
+    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
+    \midrule
+    \inputhaxx{gen-table-erdos}
+    \bottomrule
+\end{tabular}
+\caption{Experimental results for Erd\H{o}s collaboration graphs. For each
+graph, we consider $k$ equal to 2, 3 and 4. In each case we show the density of
+$G^k$, the size of a maximum $k$-clique, and then for both the unmodified
+algorithm and the algorithm with our lazy global domination step, the number of
+nodes required, and the runtime in seconds.}\label{table:erdos}
+\end{table}
 
@@
 
-In the second part of \cref{table:results} we present results from the ``clique'' graphs from the
-Second DIMACS implementation challenge\footnote{\texttt{http://dimacs.rutgers.edu/Challenges/}}. These
+In \cref{table:clique} we present results from the ``clique'' graphs from the
+Second DIMACS implementation challenge\footnote{http://dimacs.rutgers.edu/Challenges/}. These
 graphs were designed to test maximum clique implementations. Nearly all of these graphs have
@@
 ``p\_hat'' graphs.
+
+\begin{table}
+    \tiny
+    \begin{tabular}{l c rr rr rr}
+        \toprule
+        & & & & \multicolumn{2}{c}{Unmodified} & \multicolumn{2}{c}{With Domination} \\
+    \cmidrule(lr){5-6}
+    \cmidrule(lr){7-8}
+    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
+    \midrule
+    \inputhaxx{gen-table-dimacs}
+    \bottomrule
+\end{tabular}
+\caption{Experimental results for DIMACS clique graphs with diameter greater
+    than two. For each graph, we consider $k$ equal to 2, 3 and 4. In each case
+    we show the density of $G^k$, the size of a maximum $k$-clique, and then
+    for both the unmodified algorithm and the algorithm with our lazy global
+domination step, the number of nodes required, and the runtime in seconds.}\label{table:clique}
+\end{table}
 
@@
 
-The third part of \cref{table:results} presents results from the smallest 20 partitioning graphs
+\Cref{table:partitioning} presents results from the smallest 20 partitioning graphs
 from the 10th DIMACS Implementation
-Challenge\footnote{\texttt{http://staffweb.cms.gre.ac.uk/~wc06/partition/}}.  Many of these graphs are
+Challenge\footnote{http://staffweb.cms.gre.ac.uk/\textasciitilde{}wc06/partition/}.  Many of these graphs are
 considerably larger than those typically considered for the maximum clique problem, and we might
@@
 rule, there are two).
+
+\begin{table}
+    \tiny\setlength{\tabcolsep}{4pt}
+    \begin{tabular}{l c rr rr rr}
+        \toprule
+        & & & & \multicolumn{2}{c}{Unmodified} & \multicolumn{2}{c}{With Domination} \\
+    \cmidrule(lr){5-6}
+    \cmidrule(lr){7-8}
+    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
+    \midrule
+    \inputhaxx{gen-table-dimacs10walshaw}
+    \bottomrule
+\end{tabular}
+\caption{Experimental results for the smaller DIMACS partitioning graphs. For
+    each graph, we consider $k$ equal to 2, 3 and 4. In each case we show the
+    density of $G^k$, the size of a maximum $k$-clique, and then for both the
+    unmodified algorithm and the algorithm with our lazy global domination
+step, the number of nodes required, and the runtime in seconds.}\label{table:partitioning}
+\end{table}
 
@@
 
-The final part of \cref{table:results} presents results from the smallest 20 partitioning graphs
+\Cref{table:clustering} presents results from the smallest 20 clustering graphs
 from the 10th DIMACS Implementation
-Challenge\footnote{\texttt{http://www.cc.gatech.edu/dimacs10/archive/clustering.shtml}}. Again, from a
+Challenge\footnote{http://www.cc.gatech.edu/dimacs10/archive/clustering.shtml}. Again, from a
 maximum clique perspective these would be considered unusually large graphs. However, only five were
@@
 problems took under two seconds.
+
+\begin{table}
+    \tiny\setlength{\tabcolsep}{4pt}
+    \begin{tabular}{l c rr rr rr}
+        \toprule
+        & & & & \multicolumn{2}{c}{Unmodified} & \multicolumn{2}{c}{With Domination} \\
+    \cmidrule(lr){5-6}
+    \cmidrule(lr){7-8}
+    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
+    \midrule
+    \inputhaxx{gen-table-dimacs10cluster}
+    \bottomrule
+\end{tabular}
+\caption{Experimental results for the smaller DIMACS clustering graphs. For
+    each graph, we consider $k$ equal to 2, 3 and 4. In each case we show the
+    density of $G^k$, the size of a maximum $k$-clique, and then for both the
+    unmodified algorithm and the algorithm with our lazy global domination
+step, the number of nodes required, and the runtime in seconds.}\label{table:clustering}
+\end{table}
 

### 2016-03-29 paper/paper.tex
Anonymous

@@
 \setlength{\pdfpageheight}{11in}
+% \pdfinfo{
+% /Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
+% /Author (Ciaran McCreesh, Patrick Prosser)}
 \pdfinfo{
 /Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
-/Author (Ciaran McCreesh, Patrick Prosser)}
+/Author (First Author and Second Author)}
 \setcounter{secnumdepth}{0}
@@
 \title{Finding Maximum $k$-Cliques Faster using Lazy Global Domination}
+% \author{
+%    Ciaran McCreesh%
+%    \thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/K503058/1]}
+%    \and
+%    Patrick Prosser
+%    \\ University of Glasgow, Glasgow, Scotland
+%}
 \author{
-    Ciaran McCreesh%
-    \thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/K503058/1]}
+    First Author%
+    \thanks{This work was supported by a pile of unmarked bills from an anonymous funding agency}
     \and
-    Patrick Prosser
-    \\ University of Glasgow, Glasgow, Scotland
+    Second Author
+    \\ Secret Department of Secrecy, Some Institution, Somewhere
 }

### 2016-03-29 paper/paper.tex
Notation clash

@@
 A clique in a graph is a set of vertices, each of which is adjacent to every other vertex in the
-set. Finding a clique of maximum size in a graph is one of the basic \NP-hard
-problems \cite{Garey:1990}; applications include geometry, coding theory, computer vision and
-bioinformatics \cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may
-be too strong a requirement. A $k$-clique (or sometimes $n$-clique or $s$-clique) is a relaxed form
-of clique, where instead of requiring each pair of vertices to be directly adjacent, we only require
-that they be connected by a path of length at most $k$ \cite{Luce:1950}. Thus a $1$-clique \emph{is} a
-clique, a $2$-clique may be thought of as ``a group of people, all of whom either know each other or
-have a mutual acquaintance'', and so on. We illustrate this in \cref{figure:k-cliques}.
-Determining the size of a maximum $k$-clique is \NP-hard for any fixed $k$ \cite{Bourjolly:2002}.
+set. Finding a clique of maximum size in a graph is one of the basic \NP-hard problems
+\cite{Garey:1990}; applications include geometry, coding theory, computer vision and bioinformatics
+\cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may be too strong
+a requirement. A $k$-clique (or sometimes $n$-clique or $s$-clique) is a relaxed form of clique,
+where instead of requiring each pair of vertices to be directly adjacent, we only require that they
+be connected by a path of length at most $k$ \cite{Luce:1950} \footnote{In an unfortunate clash of
+notation, the name ``$k$-clique problem'' is sometimes used elsewhere to refer the decision version
+of the clique problem, to distinguish it from the maximum clique problem.}.  Thus a $1$-clique
+\emph{is} a clique, a $2$-clique may be thought of as ``a group of people, all of whom either know
+each other or have a mutual acquaintance'', and so on. We illustrate this in
+\cref{figure:k-cliques}.  Determining the size of a maximum $k$-clique is \NP-hard for any fixed $k$
+\cite{Bourjolly:2002}.
 

### 2016-04-05 code/degree_sort.cc
Some code

@@ -1,0 +1,19 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "degree_sort.hh"
+
+#include <algorithm>
+#include <iterator>
+
+auto degree_sort(const Graph & graph, std::vector<int> & p, bool reverse) -> void
+{
+    // pre-calculate degrees
+    std::vector<int> degrees;
+    std::transform(p.begin(), p.end(), std::back_inserter(degrees),
+            [&] (int v) { return graph.degree(v); });
+
+    // sort on degree
+    std::sort(p.begin(), p.end(),
+            [&] (int a, int b) { return (! reverse) ^ (degrees[a] < degrees[b] || (degrees[a] == degrees[b] && a > b)); });
+}
+

### 2016-04-05 code/dimacs.cc

@@ -1,0 +1,59 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "dimacs.hh"
+#include "graph.hh"
+#include "graph_file_error.hh"
+#include <boost/regex.hpp>
+#include <fstream>
+
+auto read_dimacs(const std::string & filename) -> Graph
+{
+    Graph result(0, true);
+
+    std::ifstream infile{ filename };
+    if (! infile)
+        throw GraphFileError{ filename, "unable to open file" };
+
+    std::string line;
+    while (std::getline(infile, line)) {
+        if (line.empty())
+            continue;
+
+        /* Lines are comments, a problem description (contains the number of
+         * vertices), or an edge. */
+        static const boost::regex
+            comment{ R"(c(\s.*)?)" },
+            problem{ R"(p\s+(edge|col)\s+(\d+)\s+(\d+)?\s*)" },
+            edge{ R"(e\s+(\d+)\s+(\d+)\s*)" };
+
+        boost::smatch match;
+        if (regex_match(line, match, comment)) {
+            /* Comment, ignore */
+        }
+        else if (regex_match(line, match, problem)) {
+            /* Problem. Specifies the size of the graph. Must happen exactly
+             * once. */
+            if (0 != result.size())
+                throw GraphFileError{ filename, "multiple 'p' lines encountered" };
+            result.resize(std::stoi(match.str(2)));
+        }
+        else if (regex_match(line, match, edge)) {
+            /* An edge. DIMACS files are 1-indexed. We assume we've already had
+             * a problem line (if not our size will be 0, so we'll throw). */
+            int a{ std::stoi(match.str(1)) }, b{ std::stoi(match.str(2)) };
+            if (0 == a || 0 == b || a > result.size() || b > result.size())
+                throw GraphFileError{ filename, "line '" + line + "' edge index out of bounds" };
+            else if (a == b)
+                throw GraphFileError{ filename, "line '" + line + "' contains a loop on vertex " + std::to_string(a) };
+            result.add_edge(a - 1, b - 1);
+        }
+        else
+            throw GraphFileError{ filename, "cannot parse line '" + line + "'" };
+    }
+
+    if (! infile.eof())
+        throw GraphFileError{ filename, "error reading file" };
+
+    return result;
+}
+

### 2016-04-05 code/graph.cc

@@ -1,0 +1,63 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph.hh"
+#include <algorithm>
+
+Graph::Graph(int size, bool add_one_for_output) :
+    _add_one_for_output(add_one_for_output)
+{
+    if (0 != size)
+        resize(size);
+}
+
+auto Graph::_position(int a, int b) const -> AdjacencyMatrix::size_type
+{
+    return (a * _size) + b;
+}
+
+auto Graph::resize(int size) -> void
+{
+    _size = size;
+    _adjacency.resize(size * size);
+}
+
+auto Graph::add_edge(int a, int b) -> void
+{
+    _adjacency[_position(a, b)] = 1;
+    _adjacency[_position(b, a)] = 1;
+}
+
+auto Graph::adjacent(int a, int b) const -> bool
+{
+    return _adjacency[_position(a, b)];
+}
+
+auto Graph::size() const -> int
+{
+    return _size;
+}
+
+auto Graph::degree(int a) const -> int
+{
+    return std::count_if(
+            _adjacency.begin() + a * _size,
+            _adjacency.begin() + (a + 1) * _size,
+            [] (int x) { return !!x; });
+}
+
+auto Graph::vertex_name(int a) const -> std::string
+{
+    if (_add_one_for_output)
+        return std::to_string(a + 1);
+    else
+        return std::to_string(a);
+}
+
+auto Graph::vertex_number(const std::string & t) const -> int
+{
+    if (_add_one_for_output)
+        return std::stoi(t) - 1;
+    else
+        return std::stoi(t);
+}
+

### 2016-04-05 code/graph_file_error.cc

@@ -1,0 +1,14 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph_file_error.hh"
+
+GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+    _what("Error reading graph file '" + filename + "': " + message)
+{
+}
+
+auto GraphFileError::what() const throw () -> const char *
+{
+    return _what.c_str();
+}
+

### 2016-04-05 code/kneighbours.cc

@@ -1,0 +1,79 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "kneighbours.hh"
+
+KNeighbours::KNeighbours(const Graph & graph, const int nk, const std::vector<int> * maybe_restrict) :
+    vertices(graph.size())
+{
+    for (auto & d : vertices) {
+        d.firsts.resize(nk + 1);
+        for (auto & f : d.firsts)
+            f = -1;
+
+        d.distances.resize(graph.size());
+    }
+
+    /* build up distance 1 lists */
+    for (int i = 0 ; i < graph.size() ; ++i) {
+        if (maybe_restrict && ! maybe_restrict->at(i))
+            continue;
+
+        int prev = -1;
+        for (int j = 0 ; j < graph.size() ; ++j) {
+            if (maybe_restrict && ! maybe_restrict->at(j))
+                continue;
+
+            if (i == j) {
+                vertices[i].firsts[0] = i;
+                vertices[i].distances[j].distance = 0;
+                vertices[i].distances[j].next = -1;
+            }
+            else if (graph.adjacent(i, j)) {
+                if (-1 == vertices[i].firsts[1]) {
+                    vertices[i].firsts[1] = j;
+                    vertices[i].distances[j].distance = 1;
+                    vertices[i].distances[j].next = -1;
+                    prev = j;
+                }
+                else {
+                    vertices[i].distances[prev].next = j;
+                    vertices[i].distances[j].distance = 1;
+                    vertices[i].distances[j].next = -1;
+                    prev = j;
+                }
+            }
+            else {
+                vertices[i].distances[j].distance = -1;
+                vertices[i].distances[j].next = -1;
+            }
+        }
+    }
+
+    /* build up distance k lists */
+    for (int k = 2 ; k <= nk ; ++k) {
+        // for each vertex i...
+        for (int i = 0 ; i < graph.size() ; ++i) {
+            if (maybe_restrict && ! maybe_restrict->at(i))
+                continue;
+
+            int prev = -1;
+            // for each vertex a at distance k - 1 from i...
+            for (int a = vertices[i].firsts[k - 1] ; a != -1 ; a = vertices[i].distances[a].next) {
+                // for each vertex j at distance 1 from a...
+                for (int j = vertices[a].firsts[1] ; j != -1 ; j = vertices[a].distances[j].next) {
+                    // are i and j currently infinitely far apart?
+                    if (vertices[i].distances[j].distance == -1) {
+                        // distance from i to j is now k, via a.
[…]

### 2016-04-05 code/max_clique_params.cc

@@ -1,0 +1,6 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <max_clique/max_clique_params.hh>
+
+using namespace parasols;
+

### 2016-04-05 code/max_clique_result.cc

@@ -1,0 +1,18 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <max_clique/max_clique_result.hh>
+
+using namespace parasols;
+
+auto MaxCliqueResult::merge(const MaxCliqueResult & other) -> void
+{
+    nodes += other.nodes;
+    donations += other.donations;
+    result_count += other.result_count;
+    result_club_count += other.result_club_count;
+    if (other.size > size) {
+        size = other.size;
+        members = std::move(other.members);
+    }
+}
+

### 2016-04-05 code/metis.cc

@@ -1,0 +1,99 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "metis.hh"
+#include "graph.hh"
+#include "graph_file_error.hh"
+#include <boost/regex.hpp>
+#include <fstream>
+
+auto read_metis(const std::string & filename) -> Graph
+{
+    Graph result(0, true);
+
+    std::ifstream infile{ filename };
+    if (! infile)
+        throw GraphFileError{ filename, "unable to open file" };
+
+    /* Lines are comments, a problem description (contains the number of
+     * vertices), or an edge. */
+    static const boost::regex
+        comment{ R"(%.*)" },
+        problem{ R"((\d+)\s+(\d+)(\s+(\d+)(\s+(\d+))?)?)" };
+
+    bool weighted_edges = false;
+    std::string line;
+    while (std::getline(infile, line)) {
+        if (line.empty())
+            continue;
+
+        boost::smatch match;
+        if (regex_match(line, match, comment)) {
+            /* comment */
+        }
+        else if (regex_match(line, match, problem)) {
+            result.resize(std::stoi(match.str(1)));
+            if (! match.str(4).empty()) {
+                if (match.str(4) == "1")
+                    weighted_edges = true;
+                else if (match.str(4) != "0")
+                    throw GraphFileError{ filename, "unsupported fmt " + match.str(4) + " is not 0 or 1" };
+            }
+
+            if (! match.str(6).empty())
+                if (match.str(6) != "0")
+                    throw GraphFileError{ filename, "unsupported ncon " + match.str(6) + " is not 0" };
+
+            break;
+        }
+        else
+            throw GraphFileError{ filename, "could not parse first line" };
+    }
+
+    if (0 == result.size())
+        throw GraphFileError{ filename, "no problem line found" };
+
+    int row = 0;
+    while (std::getline(infile, line)) {
+        boost::smatch match;
+        if (regex_match(line, match, comment)) {
+            /* comment */
+        }
+        else {
+            ++row;
+            std::stringstream line_s{ line };
+            int e;
+            while (line_s >> e) {
+                if (e > result.size() || e < 1)
+                    throw GraphFileError{ filename, "bad edge destination" };
+
+                if (e == row)
+                    throw GraphFileError{ filename, "loop detected" };
+
+                result.add_edge(row - 1, e - 1);
+                if (weighted_edges)
+                    line_s >> e;
+            }
+
[…]

### 2016-04-05 code/net.cc

@@ -1,0 +1,86 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "net.hh"
+#include "graph.hh"
+#include "graph_file_error.hh"
+#include <boost/regex.hpp>
+#include <fstream>
+
+auto read_net(const std::string & filename) -> Graph
+{
+    Graph result(0, true);
+
+    std::ifstream infile{ filename };
+    if (! infile)
+        throw GraphFileError{ filename, "unable to open file" };
+
+    std::string line;
+    while (std::getline(infile, line)) {
+        if (line.empty())
+            continue;
+
+        if (line.at(line.length() - 1) == '\r')
+            line.erase(line.length() - 1);
+
+        static const boost::regex
+            comment{ R"((%.*)?)" },
+            problem{ R"(\*\s*Vertices\s+(\d+))" },
+            description{ R"(\d+\s+".*")" },
+            arcs{ R"(\*\s*Arcslist)" },
+            edge_start{ R"(\*\s*Edgeslist)" };
+
+        boost::smatch match;
+        if (regex_match(line, match, comment)) {
+        }
+        else if (regex_match(line, match, description) || regex_match(line, match, arcs)) {
+        }
+        else if (regex_match(line, match, problem)) {
+            if (0 != result.size())
+                throw GraphFileError{ filename, "multiple '*Vertices' lines encountered" };
+            result.resize(std::stoi(match.str(1)));
+        }
+        else if (regex_match(line, match, edge_start)) {
+            break;
+        }
+        else
+            throw GraphFileError{ filename, "cannot parse line '" + line + "'" };
+    }
+
+    while (std::getline(infile, line)) {
+        if (line.empty())
+            continue;
+
+        if (line.at(line.length() - 1) == '\r')
+            line.erase(line.length() - 1);
+
+        std::stringstream line_s{ line };
+        int f, t;
+
+        if (! (line_s >> f))
+            throw GraphFileError{ filename, "cannot parse edge line '" + line + "'" };
+        --f;
+
+        if (f < 0 || f >= result.size())
+            throw GraphFileError{ filename, "invalid f value" };
+
+        while (line_s >> t) {
+            --t;
+            if (t < 0 || t >= result.size())
+                throw GraphFileError{ filename, "invalid t value " + std::to_string(t) + " (" + std::to_string(f) + ", " + std::to_string(result.size()) + ")" };
+
+            if (f == t)
[…]

### 2016-04-05 code/power.cc

@@ -1,0 +1,19 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "power.hh"
+#include "kneighbours.hh"
+#include <limits>
+
+Graph power(const Graph & graph, int n)
+{
+    KNeighbours distances(graph, n);
+    Graph result((graph));
+
+    for (int i = 0 ; i < graph.size() ; ++i)
+        for (int j = 0 ; j < graph.size() ; ++j)
+            if (i != j && distances.vertices[i].distances[j].distance > 0)
+                result.add_edge(i, j);
+
+    return result;
+}
+

### 2016-04-05 code/sequential.cc

@@ -1,0 +1,199 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph.hh"
+#include "bit_graph.hh"
+#include "max_clique_params.hh"
+#include "max_clique_result.hh"
+#include "degree_sort.hh"
+#include "template_voodoo.hh"
+
+#include <algorithm>
+
+namespace
+{
+    template <unsigned size_>
+    struct CCO
+    {
+        const Graph & original_graph;
+        FixedBitGraph<size_> graph;
+        const MaxCliqueParams & params;
+        std::vector<int> order;
+
+        MaxCliqueResult result;
+
+        std::list<std::set<int> > previouses;
+
+        std::vector<std::pair<bool, FixedBitSet<size_> > > unsets;
+
+        CCO(const Graph & g, const MaxCliqueParams & p) :
+            original_graph(g),
+            params(p),
+            order(g.size()),
+            unsets(g.size())
+        {
+            // populate our order with every vertex initially
+            std::iota(order.begin(), order.end(), 0);
+            degree_sort(g, order, false);
+
+            // re-encode graph as a bit graph
+            graph.resize(g.size());
+
+            for (int i = 0 ; i < g.size() ; ++i)
+                for (int j = 0 ; j < g.size() ; ++j)
+                    if (g.adjacent(order[i], order[j]))
+                        graph.add_edge(i, j);
+        }
+
+        auto expand(
+                std::vector<unsigned> & c,
+                FixedBitSet<size_> & p,
+                const std::array<BitWord, size_ * bits_per_word> & p_order,
+                const std::array<BitWord, size_ * bits_per_word> & colours
+                ) -> void
+        {
+            ++result.nodes;
+
+            int previous_v = -1;
+
+            // for each v in p... (v comes later)
+            for (int n = p.popcount() - 1 ; n >= 0 ; --n) {
+                // bound, timeout or early exit?
+                unsigned best_anywhere_value = result.size;
+                if (c.size() + colours[n] <= best_anywhere_value || params.abort->load())
+                    return;
+
+                if (params.lgd && -1 != previous_v)
+                    propagate_lazy_global_domination(previous_v, p);
+
+                auto v = p_order[n];
+                previous_v = v;
+
+                if (skip(v, p)) {
+                    p.unset(v);
+                }
+                else {
+                    // consider taking v
+                    c.push_back(v);
+
+                    // filter p to contain vertices adjacent to v
[…]

### 2016-04-05 code/solve_max_clique.cc

@@ -1,0 +1,228 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph.hh"
+#include "sequential.hh"
+#include "file_formats.hh"
+#include "power.hh"
+
+#include <boost/program_options.hpp>
+#include <boost/algorithm/string.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+namespace
+{
+    template <typename Result_, typename Params_, typename Data_>
+    auto run_this_wrapped(const std::function<Result_ (const Data_ &, const Params_ &)> & func)
+        -> std::function<Result_ (const Data_ &, Params_ &, bool &, int)>
+    {
+        return [func] (const Data_ & data, Params_ & params, bool & aborted, int timeout) -> Result_ {
+            /* For a timeout, we use a thread and a timed CV. We also wake the
+             * CV up if we're done, so the timeout thread can terminate. */
+            std::thread timeout_thread;
+            std::mutex timeout_mutex;
+            std::condition_variable timeout_cv;
+            std::atomic<bool> abort;
+            abort.store(false);
+            params.abort = &abort;
+            if (0 != timeout) {
+                timeout_thread = std::thread([&] {
+                        auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(timeout);
+                        {
+                            /* Sleep until either we've reached the time limit,
+                             * or we've finished all the work. */
+                            std::unique_lock<std::mutex> guard(timeout_mutex);
+                            while (! abort.load()) {
+                                if (std::cv_status::timeout == timeout_cv.wait_until(guard, abort_time)) {
+                                    /* We've woken up, and it's due to a timeout. */
+                                    aborted = true;
+                                    break;
+                                }
+                            }
+                        }
+                        abort.store(true);
+                        });
+            }
+
+            /* Start the clock */
+            params.start_time = std::chrono::steady_clock::now();
+            auto result = func(data, params);
+
+            /* Clean up the timeout thread */
+            if (timeout_thread.joinable()) {
+                {
+                    std::unique_lock<std::mutex> guard(timeout_mutex);
+                    abort.store(true);
+                    timeout_cv.notify_all();
+                }
+                timeout_thread.join();
+            }
+
[…]

### 2016-04-05 code/about_graph.cc
About graph

@@ -1,0 +1,124 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph.hh"
+#include "file_formats.hh"
+#include "power.hh"
+
+#include <boost/program_options.hpp>
+#include <boost/algorithm/string.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+
+namespace po = boost::program_options;
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                 "Display help information")
+            ("power",              po::value<int>(), "Raise the graph to this power")
+            ("format",             po::value<std::string>(), "Specify the format of the input")
+            ;
+
+        po::options_description all_options{ "All options" };
+        all_options.add_options()
+            ("input-file", po::value<std::vector<std::string> >(),
+                           "Specify the input file (DIMACS format, unless --format is specified). May be specified multiple times.")
+            ;
+
+        all_options.add(display_options);
+
+        po::positional_options_description positional_options;
+        positional_options
+            .add("input-file", -1)
+            ;
+
+        po::variables_map options_vars;
+        po::store(po::command_line_parser(argc, argv)
+                .options(all_options)
+                .positional(positional_options)
+                .run(), options_vars);
+        po::notify(options_vars);
+
+        /* --help? Show a message, and exit. */
+        if (options_vars.count("help")) {
+            std::cout << "Usage: " << argv[0] << " [options] file[...]" << std::endl;
+            std::cout << std::endl;
+            std::cout << display_options << std::endl;
+            return EXIT_SUCCESS;
+        }
+
+        /* No algorithm or no input file specified? Show a message and exit. */
+        if (options_vars.count("input-file") < 1) {
+            std::cout << "Usage: " << argv[0] << " [options] file[...]" << std::endl;
+            return EXIT_FAILURE;
+        }
+
+        /* For each input file... */
+        auto input_files = options_vars["input-file"].as<std::vector<std::string> >();
+        bool first = true;
+        for (auto & input_file : input_files) {
+            if (first)
+                first = false;
+            else
+                std::cout << "--" << std::endl;
+
+            /* Turn a format name into a runnable function. */
[…]

### 2016-04-05 paper/paper.tex
All hail eggy

@@
 Here we give experimental results on a range of standard benchmarks, and on real-world and random
-graphs. Where timing results are reported, the experiments were run on a machine with Intel E5645
-processors, and single-threaded runtimes are given. The time taken to read in the graph from a file
-is excluded, but preprocessing time (including the construction of $G^k$ and the bitset encoding) is
-included. We use the term \emph{nodes} to refer to the number of recursive calls made by the
-branch-and-bound part of the algorithm.
+graphs. Where timing results are reported, the experiments were run on a machine with Intel E5-2687W
+v3 processors, and single-threaded runtimes are given. The time taken to read in the graph from a
+file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
+encoding) is included. We use the term \emph{nodes} to refer to the number of recursive calls made
+by the branch-and-bound part of the algorithm.
 
@@
 domination rule. However, using the unmodified maximum clique algorithm, two of these results did
-not finish running within one day. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.
+not finish running within one hour. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.
 
@@
 With the domination rule, we solve all of these problems within a tenth of a second. Without, two of
-the results take over a day, and the rest remain trivial. Note that in several cases, for some
+the results take over an hour, and the rest remain trivial. Note that in several cases, for some
 values of $k$ a $k$-clique covers the entire graph.  Again using Chang et al.'s results
@@
 expect our $O(|V|^2)$ memory requirements to cause problems. Nonetheless, with the domination rule
-there is only one instance which we were unable to solve within a day (and without the domination
+there is only one instance which we were unable to solve within an hour (and without the domination
 rule, there are two).
@@
 maximum clique perspective these would be considered unusually large graphs. However, only five were
-unsolvable within a day (plus a sixth when the domination rule was not used), and half of the
+unsolvable within an hour (plus a sixth when the domination rule was not used), and half of the
 problems took under two seconds.

### 2016-04-06 code/max_clique_result.cc
Parallel search

@@ -1,15 +1,10 @@
 /* vim: set sw=4 sts=4 et foldmethod=syntax : */
 
-#include <max_clique/max_clique_result.hh>
-
-using namespace parasols;
+#include "max_clique_result.hh"
 
 auto MaxCliqueResult::merge(const MaxCliqueResult & other) -> void
 {
     nodes += other.nodes;
-    donations += other.donations;
-    result_count += other.result_count;
-    result_club_count += other.result_club_count;
     if (other.size > size) {
         size = other.size;
         members = std::move(other.members);

### 2016-04-06 code/parallel.cc

@@ -1,0 +1,501 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph.hh"
+#include "bit_graph.hh"
+#include "max_clique_params.hh"
+#include "max_clique_result.hh"
+#include "degree_sort.hh"
+#include "template_voodoo.hh"
+#include "queue.hh"
+
+#include <algorithm>
+#include <mutex>
+#include <condition_variable>
+#include <thread>
+#include <chrono>
+
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+
+namespace
+{
+    const constexpr int number_of_depths = 5;
+    const constexpr int number_of_steal_points = number_of_depths - 1;
+
+    struct Subproblem
+    {
+        std::vector<int> offsets;
+    };
+
+    struct QueueItem
+    {
+        Subproblem subproblem;
+    };
+
+    struct StealPoint
+    {
+        std::mutex mutex;
+        std::condition_variable cv;
+
+        bool is_finished;
+
+        bool has_data;
+        std::vector<int> data;
+
+        bool was_stolen;
+
+        StealPoint() :
+            is_finished(false),
+            has_data(false),
+            was_stolen(false)
+        {
+            mutex.lock();
+        }
+
+        void publish(std::vector<int> & s)
+        {
+            if (is_finished)
+                return;
+
+            data = s;
+            has_data = true;
+            cv.notify_all();
+            mutex.unlock();
+        }
+
+        bool steal(std::vector<int> & s) __attribute__((noinline))
+        {
+            std::unique_lock<std::mutex> guard(mutex);
+
+            while ((! has_data) && (! is_finished))
+                cv.wait(guard);
+
+            if (! is_finished && has_data) {
+                s = data;
+                was_stolen = true;
+                return true;
+            }
+            else
+                return false;
+        }
+
+        bool unpublish_and_keep_going()
+        {
+            if (is_finished)
+                return true;
+
+            mutex.lock();
+            has_data = false;
+            return ! was_stolen;
+        }
+
+        void finished()
+        {
+            is_finished = true;
+            has_data = false;
+            cv.notify_all();
+            mutex.unlock();
+        }
+    };
+
+    struct alignas(16) StealPoints
+    {
+        std::vector<StealPoint> points;
+
+        StealPoints() :
+            points{ number_of_steal_points }
+        {
+        }
+    };
+
[…]

### 2016-04-06 code/solve_max_clique.cc

@@ -2,6 +2,7 @@
 
 #include "graph.hh"
 #include "sequential.hh"
+#include "parallel.hh"
 #include "file_formats.hh"
 #include "power.hh"
 
@@ -107,6 +108,7 @@ auto main(int argc, char * argv[]) -> int
             ("power",              po::value<int>(), "Raise the graph to this power (to solve s-clique)")
             ("format",             po::value<std::string>(), "Specify the format of the input")
             ("lgd",                                  "Use lazy global domination")
+            ("parallel-search",                      "Use threaded search")
             ;
 
         po::options_description all_options{ "All options" };
@@ -181,7 +183,7 @@ auto main(int argc, char * argv[]) -> int
 
             /* Do the actual run. */
             bool aborted = false;
-            auto result = run_with_modifications(cco_max_clique)(
+            auto result = run_with_modifications(options_vars.count("parallel-search") ? tcco_max_clique : cco_max_clique)(
                         graph,
                         params,
                         aborted,

### 2016-04-06 paper/paper.tex
Parallel experiments

@@
     of each other, rather than directly adjacent. In theory, a maximum clique algorithm can easily
-    be adapted to solve the maximum $k$-clique problem. We use a state of the art maximum clique
-    algorithm to show that this is feasible in practice, and introduce a lazy global domination rule
-    which sometimes vastly reduces the search space. We include experimental results for a range of
-    real-world and benchmark graphs, and a detailed look at random graphs.
+    be adapted to solve the maximum $k$-clique problem, although large sparse $k$-clique graphs
+    reduce to large dense clique graphs, which can be computationally challenging. We adapt a state of
+    the art maximum clique algorithm to show that this reduction is in fact useful in practice, and
+    introduce a lazy global domination rule which sometimes vastly reduces the search space. We
+    include experimental results for a range of real-world and benchmark graphs, and a detailed look
+    at random graphs.
 \end{abstract}
@@
 \cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may be too strong
-a requirement. A $k$-clique (or sometimes $n$-clique or $s$-clique) is a relaxed form of clique,
-where instead of requiring each pair of vertices to be directly adjacent, we only require that they
-be connected by a path of length at most $k$ \cite{Luce:1950} \footnote{In an unfortunate clash of
-notation, the name ``$k$-clique problem'' is sometimes used elsewhere to refer the decision version
-of the clique problem, to distinguish it from the maximum clique problem.}.  Thus a $1$-clique
-\emph{is} a clique, a $2$-clique may be thought of as ``a group of people, all of whom either know
-each other or have a mutual acquaintance'', and so on. We illustrate this in
-\cref{figure:k-cliques}.  Determining the size of a maximum $k$-clique is \NP-hard for any fixed $k$
-\cite{Bourjolly:2002}.
+a requirement. A $k$-clique\footnote{Or sometimes $n$-clique or $s$-clique; in an unfortunate clash
+of notation, ``$k$-clique problem'' is sometimes used elsewhere for the decision version of the
+clique problem, to distinguish it from the maximum clique problem.} is a relaxed form of clique,
+where instead of requiring each pair of vertices to be directly adjacent, we only require they be
+connected by a path of length at most $k$ \cite{Luce:1950}.  Thus a $1$-clique \emph{is} a clique, a
+$2$-clique may be thought of as ``a group of people, all of whom either know each other or have a
+mutual acquaintance'', and so on. We illustrate this in \cref{figure:k-cliques}.  Determining the
+size of a maximum $k$-clique is \NP-hard for any fixed $k$ \cite{Bourjolly:2002}.
 
@@
 Here we give experimental results on a range of standard benchmarks, and on real-world and random
-graphs. Where timing results are reported, the experiments were run on a machine with Intel E5-2687W
-v3 processors, and single-threaded runtimes are given. The time taken to read in the graph from a
-file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
-encoding) is included. We use the term \emph{nodes} to refer to the number of recursive calls made
-by the branch-and-bound part of the algorithm.
+graphs. Where timing results are reported, the experiments were run on a machine with dual Intel
+E5-2687W v3 processors; single-threaded runtimes are given, except in \cref{table:parallel}, where
+20 threads were used (our machine does not have hyper-threading enabled). The time taken to read in
+the graph from a file is excluded, but preprocessing time (including the construction of $G^k$ and
+the bitset encoding) is included. We use the term \emph{nodes} to refer to the number of recursive
+calls made by the branch-and-bound part of the algorithm.
 
@@
 
+\subsection{Parallel Search}
+
+\begin{table}
+    \tiny\setlength{\tabcolsep}{4pt}
+    \begin{tabular}{l c rr rr rr}
+        \toprule
+        & & & & \multicolumn{2}{c}{Sequential} & \multicolumn{2}{c}{Parallel} \\
+    \cmidrule(lr){5-6}
+    \cmidrule(lr){7-8}
+    Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} &
+    \multicolumn{1}{c}{Nodes} & \multicolumn{1}{c}{Time} \\
+    \midrule
+    \inputhaxx{gen-table-parallel}
+    \bottomrule
+\end{tabular}
+\caption{Experimental results for multi-threaded search on harder instances, using 20 threads. For
+    each graph, we consider $k$ equal to 2, 3 and 4. In each case we show the
+    density of $G^k$, the size of a maximum $k$-clique, and then for both the
+    sequential algorithm and the parallel algorithm (with lazy global domination in both cases),
+    the number of nodes required, and the runtime in seconds.}\label{table:parallel}
+\end{table}
+
 \subsection{Random Graphs}

### 2016-04-06 paper/paper.tex
Tweaks

@@
     include experimental results for a range of real-world and benchmark graphs, and a detailed look
-    at random graphs.
+    at random graphs. We also use thread-parallel search to solve some harder instances.
 \end{abstract}
@@
 graphs. Where timing results are reported, the experiments were run on a machine with dual Intel
-E5-2687W v3 processors; single-threaded runtimes are given, except in \cref{table:parallel}, where
-20 threads were used (our machine does not have hyper-threading enabled). The time taken to read in
-the graph from a file is excluded, but preprocessing time (including the construction of $G^k$ and
-the bitset encoding) is included. We use the term \emph{nodes} to refer to the number of recursive
-calls made by the branch-and-bound part of the algorithm.
+E5-2687W v3 processors and 256GBytes RAM; single-threaded runtimes are given, except in
+\cref{table:parallel}, where 20 threads were used (our machine does not have hyper-threading
+enabled). Our software was implemented in C++, using C++11 native threads, and was compiled using
+GCC 5.1.0. The time taken to read in the graph from a file is excluded, but preprocessing time
+(including the construction of $G^k$ and the bitset encoding) is included. We use the term
+\emph{nodes} to refer to the number of recursive calls made by the branch-and-bound part of the
+algorithm.
 
@@
 \begin{table}
-    \tiny
+    \tiny\setlength{\tabcolsep}{6pt} % carefully tuned!
+    \centering
     \begin{tabular}{l c rr rr rr}
@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{4pt}
+    \tiny\setlength{\tabcolsep}{4pt} % carefully tuned!
+    \centering
     \begin{tabular}{l c rr rr rr}
@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{4pt}
+    \tiny\setlength{\tabcolsep}{5pt} % carefully tuned!
+    \centering
     \begin{tabular}{l c rr rr rr}
@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{4pt}
+    \tiny\setlength{\tabcolsep}{4.6pt} % carefully tuned!
+    \centering
     \begin{tabular}{l c rr rr rr}

### 2016-04-07 code/parallel.cc
Snoopy

@@ -13,6 +13,7 @@
 #include <condition_variable>
 #include <thread>
 #include <chrono>
+#include <iostream>
 
 using std::chrono::milliseconds;
 using std::chrono::steady_clock;
@@ -313,6 +314,12 @@ namespace
                 colour_class_order(initial_p, initial_p_order, initial_colours);
             }
 
+            // snoopy
+            std::mutex snoop_mutex;
+            std::condition_variable snoop_cv;
+            bool snoops_changed = false, finished = false;
+            std::vector<Subproblem> snoops(std::thread::hardware_concurrency());
+
             /* workers */
             for (unsigned i = 0 ; i < std::thread::hardware_concurrency() ; ++i) {
                 threads.push_back(std::thread([&, i] {
@@ -356,6 +363,12 @@ namespace
                                     position.reserve(graph.size());
                                     position.push_back(0);
 
+                                    {
+                                        std::unique_lock<std::mutex> guard(snoop_mutex);
+                                        snoops.at(i) = args.subproblem;
+                                        snoops_changed = true;
+                                    }
+
                                     // do some work
                                     expand(c, p, initial_p_order, initial_colours, position, local_result,
                                             &args.subproblem, &thread_steal_points.at(i));
@@ -377,9 +390,37 @@ namespace
                             }));
             }
 
+            // snoop thread
+            std::thread snoop_thread([&] {
+                    while (! finished) {
+                        auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(1);
+                        std::unique_lock<std::mutex> guard(snoop_mutex);
+                        if (std::cv_status::timeout == snoop_cv.wait_until(guard, abort_time) && snoops_changed) {
+                            snoops_changed = false;
+                            std::cerr << "active:";
+                            for (auto & s : snoops) {
+                                std::cerr << " ( ";
+                                for (auto & p : s.offsets)
+                                    std::cerr << p << " ";
+                                std::cerr << ")";
+                            }
+                            std::cerr << " bound " << get_best_anywhere_value();
+                            std::cerr << std::endl;
+                        }
+                    }
+                    });
+
             // wait until they're done, and clean up threads
             for (auto & t : threads)
                 t.join();
+
+            // kill the snooper
+            {
+                std::unique_lock<std::mutex> guard(snoop_mutex);
+                finished = true;
+                snoop_cv.notify_all();
+            }
+            snoop_thread.join();
 
             return global_result;
         }

### 2016-04-07 code/parallel.cc
Less snoopy

@@ -393,7 +393,7 @@ namespace
             // snoop thread
             std::thread snoop_thread([&] {
                     while (! finished) {
-                        auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(1);
+                        auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(60);
                         std::unique_lock<std::mutex> guard(snoop_mutex);
                         if (std::cv_status::timeout == snoop_cv.wait_until(guard, abort_time) && snoops_changed) {
                             snoops_changed = false;

### 2016-04-07 code/parallel.cc
Record position properly

@@ -199,6 +199,8 @@ namespace
 
             // for each v in p... (v comes later)
             for (int n = p.popcount() - 1 ; n >= 0 ; --n) {
+                ++position.back();
+
                 // bound, timeout or early exit?
                 unsigned best_anywhere_value = get_best_anywhere_value();
                 if (c.size() + colours[n] <= best_anywhere_value || params.abort->load())

### 2016-04-07 code/parallel.cc
A bit more parallely goodness

@@ -21,7 +21,7 @@ using std::chrono::duration_cast;
 
 namespace
 {
-    const constexpr int number_of_depths = 5;
+    const constexpr int number_of_depths = 6;
     const constexpr int number_of_steal_points = number_of_depths - 1;
 
     struct Subproblem
@@ -381,6 +381,12 @@ namespace
 
                                 if (depth < number_of_steal_points)
                                     thread_steal_points.at(i).points.at(depth).finished();
+                            }
+
+                            {
+                                std::unique_lock<std::mutex> guard(snoop_mutex);
+                                snoops.at(i).offsets = { -1 };
+                                snoops_changed = true;
                             }
 
                             // merge results

### 2016-04-07 code/parallel.cc
We need to go deeper

@@ -21,7 +21,7 @@ using std::chrono::duration_cast;
 
 namespace
 {
-    const constexpr int number_of_depths = 6;
+    const constexpr int number_of_depths = 8;
     const constexpr int number_of_steal_points = number_of_depths - 1;
 
     struct Subproblem
@@ -320,7 +320,7 @@ namespace
             std::mutex snoop_mutex;
             std::condition_variable snoop_cv;
             bool snoops_changed = false, finished = false;
-            std::vector<Subproblem> snoops(std::thread::hardware_concurrency());
+            std::vector<std::pair<Subproblem, bool> > snoops(std::thread::hardware_concurrency());
 
             /* workers */
             for (unsigned i = 0 ; i < std::thread::hardware_concurrency() ; ++i) {
@@ -367,13 +367,19 @@ namespace
 
                                     {
                                         std::unique_lock<std::mutex> guard(snoop_mutex);
-                                        snoops.at(i) = args.subproblem;
+                                        snoops.at(i) = { args.subproblem, true };
                                         snoops_changed = true;
                                     }
 
                                     // do some work
                                     expand(c, p, initial_p_order, initial_colours, position, local_result,
                                             &args.subproblem, &thread_steal_points.at(i));
+
+                                    {
+                                        std::unique_lock<std::mutex> guard(snoop_mutex);
+                                        snoops.at(i) = { args.subproblem, false };
+                                        snoops_changed = true;
+                                    }
 
                                     // record the last time we finished doing useful stuff
                                     overall_time = duration_cast<milliseconds>(steady_clock::now() - start_time);
@@ -385,7 +391,7 @@ namespace
 
                             {
                                 std::unique_lock<std::mutex> guard(snoop_mutex);
-                                snoops.at(i).offsets = { -1 };
+                                snoops.at(i) = { { { -1 } }, false };
                                 snoops_changed = true;
                             }
 
@@ -401,19 +407,26 @@ namespace
             // snoop thread
             std::thread snoop_thread([&] {
                     while (! finished) {
+                        auto old_bound_value = get_best_anywhere_value();
                         auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(60);
                         std::unique_lock<std::mutex> guard(snoop_mutex);
-                        if (std::cv_status::timeout == snoop_cv.wait_until(guard, abort_time) && snoops_changed) {
-                            snoops_changed = false;
-                            std::cerr << "active:";
[…]

### 2016-04-07 code/parallel.cc
Ideas

@@ -203,8 +203,10 @@ namespace
 
                 // bound, timeout or early exit?
                 unsigned best_anywhere_value = get_best_anywhere_value();
-                if (c.size() + colours[n] <= best_anywhere_value || params.abort->load())
+                if (c.size() + colours[n] <= best_anywhere_value || params.abort->load()) {
+                    // kill_queue_items_to_the_right_of(position);
                     return;
+                }
 
                 if (params.lgd && -1 != previous_v)
                     propagate_lazy_global_domination(previous_v, p);
@@ -342,10 +344,16 @@ namespace
                                             stole.pop_back();
                                             for (auto & s : stole)
                                                 --s;
+
+                                            // queues[depth]->enqueue_range(QueueItem{ Subproblem{ stole } }, graph.size());
+
                                             while (++stole.back() < graph.size())
                                                 queues[depth]->enqueue(QueueItem{ Subproblem{ stole } });
                                         }
                                     }
+
+                                    // queues[depth]->sort();
+
                                     queues[depth]->initial_producer_done();
                                 }
 

### 2016-04-08 code/parallel.cc
Deeeeeeper

@@ -21,7 +21,7 @@ using std::chrono::duration_cast;
 
 namespace
 {
-    const constexpr int number_of_depths = 8;
+    const constexpr int number_of_depths = 16;
     const constexpr int number_of_steal_points = number_of_depths - 1;
 
     struct Subproblem

### 2016-04-08 paper/paper.tex
More words

@@
 \usepackage{booktabs}                  % less ugly tables
-\usepackage{tikz}                      % For pretty pictures
-\usepackage{afterpage}
 \usepackage[tight-spacing=true]{siunitx}
@@
 
-A recent survey by Shahinpour and Butenko discusses algorithms and results for $k$-clique and
-$k$-club problems \cite{Shahinpour:2013}.  We adopt their notation of $\tilde{\omega}_k$ for the
-size of a maximum $k$-clique; the use of $\omega$ for the size of a maximum clique is standard.
+A recent survey by \citeauthor{Shahinpour:2013} \shortcite{Shahinpour:2013} discusses algorithms and
+results for $k$-clique and $k$-club problems .  We adopt their notation of $\tilde{\omega}_k$ for
+the size of a maximum $k$-clique; the use of $\omega$ for the size of a maximum clique is standard.
 They note that ``unlike the maximum clique problem, the maximum $s$-clique problem has not been the
@@
 are available
-\cite{Bourjolly:2000,Bourjolly:2002,Mahdavi:2012,Hartung:2012,Chang:2013,Shahinpour:2013,Wotzlaw:2014}.
+\cite{Bourjolly:2000,Bourjolly:2002,Mahdavi:2012,Hartung:2012,Chang:2013,Shahinpour:2013,Wotzlaw:2014,Picker:2015,Carvalho:2016}.
 
@@
 graphs can be very challenging computationally. Here we take a state-of-the-art maximum clique
-algorithm which is suitable for use on dense graphs \cite{Prosser:2012,McCreesh:2013}, and
+algorithm which is suitable for use on dense graphs \cite{Prosser:2012}, and
 investigate whether this approach is feasible in practice. We modify the algorithm to include a new
@@
 a path of length at most $k$ between $v_1$ and $v_2$ in $G$. We may construct this graph using a
-bounded breadth-first search: we refer to Chang et al.\ \cite{Chang:2013} for how to implement this
-quickly in practice. Now it is easy to see that maximum cliques in $G^k$ correspond with maximum
-$k$-cliques in $G$ \cite{Balasundaram:2005}.
+bounded breadth-first search: we refer to \citeauthor{Chang:2013} \shortcite{Chang:2013} for how to
+implement this quickly in practice. Now it is easy to see that maximum cliques in $G^k$ correspond
+with maximum $k$-cliques in $G$ \cite{Balasundaram:2005}.
 
@@
 The current state-of-the-art for the maximum clique problem on dense graphs, due to Tomita et al.\
-\cite{Tomita:2003,Tomita:2007,Tomita:2010}, is to use branch and bound with a greedy graph
+\cite{Tomita:2007,Tomita:2010}, is to use branch and bound with a greedy graph
 colouring.  A \emph{colouring} of a graph is an assignment of colours to vertices, such that
@@
 
-The key step in Tomita's algorithms is to produce a constructive colouring, which is used in a
-clever way.  The $\colourOrder$ routine does not just return the number of colours used. Instead, it
-returns a pair of arrays, $\order$ and $\bounds$. The $\order$ array contains vertices, in the order
-in which they were coloured. The $i$th entry of the $\bounds$ array contains the colour number used
-for the $i$th vertex in $\order$. We illustrate this in \cref{figure:colours}. Crucially, $\bounds$
-is non-decreasing (i.e.\ $\bounds[i+1] \ge \bounds[i]$), and we may colour the subgraph induced by
-the first $i$ vertices of $\order$ using $\bounds[i]$ colours.
+The key step in \citeauthor{Tomita:2010}'s algorithms is to produce a constructive colouring, which
+is used in a clever way.  The $\colourOrder$ routine does not just return the number of colours
+used. Instead, it returns a pair of arrays, $\order$ and $\bounds$. The $\order$ array contains
+vertices, in the order in which they were coloured. The $i$th entry of the $\bounds$ array contains
+the colour number used for the $i$th vertex in $\order$. We illustrate this in
+\cref{figure:colours}. Crucially, $\bounds$ is non-decreasing (i.e.\ $\bounds[i+1] \ge \bounds[i]$),
+and we may colour the subgraph induced by the first $i$ vertices of $\order$ using $\bounds[i]$
+colours.
 
@@
 The order in which vertices are selected for colouring can have a large effect upon performance.
-Various initial vertex orderings have been considered for the maximum clique problem---we refer to a
-computational study by Prosser for details \cite{Prosser:2012}. Here we will colour vertices in a
-static non-increasing degree order, which we do by permuting the graph at the top of search
-(\mcline{permute}). We will \emph{not} be using Tomita et al.'s dynamic tie-breaking mechanism
-\cite{Tomita:2010}: although doing so can sometimes be beneficial for small dense graphs in a
-maximum clique context, for the larger graphs we will be considering here the cubic cost is
-prohibitively expensive. For the same reason, we use a simple greedy colouring and do not use Tomita
-et al.'s (cubic) colour repair step \cite{Tomita:2010}.
+Various initial vertex orderings have been considered for the maximum clique problem
+\cite{Prosser:2012,DBLP:conf/lion/SegundoLB14}. Here we will colour vertices in a static
+non-increasing degree order, which we do by permuting the graph at the top of search
+(\mcline{permute}). We will \emph{not} be using a dynamic tie-breaking mechanism: although doing so
+can sometimes be beneficial for small dense graphs in a maximum clique context \cite{Tomita:2010},
+for the larger graphs we will be considering here the cubic cost is prohibitively expensive. For the
+same reason, and additionally because our colour classes typically contain many vertices, we use a
+simple greedy colouring and do not use a colour repair step \cite{Tomita:2010} or stronger
+MaxSAT-based inference \cite{DBLP:journals/cor/SegundoNB15,DBLP:conf/lion/LiJX15}.
 
@@
 complicated initial vertex orderings, and cubic colourings). If we ignore the lines shown in blue,
-we obtain the maximum clique algorithm variation that Prosser \cite{Prosser:2012} calls ``MCSa1''.
-Now we will introduce a new lazy global domination rule which performs additional inference during
-search. This rule is not specific to the maximum $k$-clique problem, and is also valid for the
-maximum clique problem.
+we obtain the maximum clique algorithm variation that \citeauthor{Prosser:2012}
+\shortcite{Prosser:2012} calls ``MCSa1''.  Now we will introduce a new lazy global domination rule
+which performs additional inference during search. This rule is not specific to the maximum
+$k$-clique problem, and is also valid for the maximum clique problem.
 
@@
 
-\paragraph{Bitset encoding} San Segundo et al.\ \cite{SanSegundo:2011,SanSegundo:2013} observed that
-the performance of Tomita's algorithms could be enhanced substantially by using a bitset encoding to
-obtain a form of SIMD-like parallelism, without altering the steps taken. We have taken such an
-approach here too, although we do not describe it explicitly---when permuting $G$ on line
+\paragraph{Bitset encoding} San Segundo et al.\ \shortcite{SanSegundo:2011,SanSegundo:2013} observed
+that the performance of Tomita's algorithms could be enhanced substantially by using a bitset
+encoding to obtain a form of SIMD-like parallelism, without altering the steps taken. We have taken
+such an approach here too, although we do not describe it explicitly---when permuting $G$ on line
 \mcline{permute}, the graph should be re-encoded as an array of adjacency bitsets. (It is not
@@
 
-Here we give experimental results on a range of standard benchmarks, and on real-world and random
-graphs. Where timing results are reported, the experiments were run on a machine with dual Intel
-E5-2687W v3 processors and 256GBytes RAM; single-threaded runtimes are given, except in
-\cref{table:parallel}, where 20 threads were used (our machine does not have hyper-threading
-enabled). Our software was implemented in C++, using C++11 native threads, and was compiled using
-GCC 5.1.0. The time taken to read in the graph from a file is excluded, but preprocessing time
-(including the construction of $G^k$ and the bitset encoding) is included. We use the term
-\emph{nodes} to refer to the number of recursive calls made by the branch-and-bound part of the
-algorithm.
+We now give experimental results on a range of standard benchmarks, and on real-world and random
+graphs. Experiments were run on a machine with dual Intel E5-2687W v3 processors and 256GBytes
+RAM; single-threaded runtimes are given, except in \cref{table:parallel}, where 20 threads were used
+(our machine does not have hyper-threading enabled). Our software was implemented in C++, using
+C++11 native threads, and was compiled using GCC 5.1.0. The time taken to read in the graph from a
+file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
+encoding) is included. We use the term \emph{nodes} to refer to the number of recursive calls made
+by the branch-and-bound part of the algorithm.
 
@@
 We begin with a selection of real-world and standard benchmark graphs. We look at $k$ equal to 2, 3
-and 4 in every case---this is a standard practice for the $k$-club problem
-\cite{Chang:2013,Wotzlaw:2014}.
+and 4 in every case---this is a standard practice for the $k$-club problem.
 
@@
 Mrvar\footnote{http://vlado.fmf.uni-lj.si/pub/networks/data/}.  We were able to solve all of
-these problems in under eight minutes (and all but three in under two seconds) when using the
-domination rule. However, using the unmodified maximum clique algorithm, two of these results did
+these problems in under four minutes (and all but three in under four seconds) when using the
+domination rule. However, using the unmodified maximum clique algorithm, three of these results did
 not finish running within one hour. Note that for $k = 4$, a $k$-clique covers all of ``Erdos02''.
@@
 
-By comparing these results with the $k$-club results of Chang et al.\ \cite{Chang:2013}, we see that
+By comparing these results with the $k$-club results of Chang et al.\ \shortcite{Chang:2013}, we see that
 in all but four cases the $k$-clique and $k$-club numbers are equal; all of these differences occur
 when $k = 4$. (Chang et al.\ did not investigate the ``Erdos02'' graph, but Wotzlaw
-\cite{Wotzlaw:2014} confirmed privately that the $k$-clique and $k$-club numbers are the same here
+\shortcite{Wotzlaw:2014} confirmed privately that the $k$-clique and $k$-club numbers are the same here
 too.) On the other hand, the $k$-clique numbers are sometimes much easier to find, both
@@
 the results take over an hour, and the rest remain trivial. Note that in several cases, for some
-values of $k$ a $k$-clique covers the entire graph.  Again using Chang et al.'s results
-\cite{Chang:2013}, we see that for the first six graphs in this table the $k$-clique and $k$-club
+values of $k$ a $k$-clique covers the entire graph.  Again using Chang et al.'s results, we see that for the first six graphs in this table the $k$-clique and $k$-club
 numbers are the same for each value of $k$ (Chang et al.\ did not investigate ``c-fat500-10'' or
@@
 where the proof of optimality is not immediate: in ``3elt'' and ``4elt'', our runtimes can nearly
-double, and for ``cs4'' and ``cti'' the slowdown is sometimes over a factor of ten. In other words,
-laziness does not help when the rule turns out to be used, but useless.
+double, and for ``cs4'' and ``cti'' the slowdown is sometimes over a factor of ten. Thus laziness
+can be costly when the rule is used, but useless.
 
-Five of these graphs were considered for the $k$-club problem by Wotzlaw \cite{Wotzlaw:2014}. In all
-five cases, the $k$-clique and $k$-club numbers are the same for $k$ equal to 2, 3 and 4. However,
-the $k$-clique number was again consistently much easier to find.
+Five of these graphs were considered for the $k$-club problem by Wotzlaw \shortcite{Wotzlaw:2014}.
+In every case, the $k$-clique and $k$-club numbers are the same. However, the $k$-clique number was
+again consistently much easier to find.
 
@@
 maximum clique perspective these would be considered unusually large graphs. However, only five were
-unsolvable within an hour (plus a sixth when the domination rule was not used), and half of the
+unsolvable within an hour (plus a sixth when the domination rule was not used), and over half of the
 problems took under two seconds.
@@
 
-Seven of these graphs were considered for the $k$-club problem by Wotzlaw \cite{Wotzlaw:2014}. In
+Seven of these graphs were considered for the $k$-club problem by Wotzlaw \shortcite{Wotzlaw:2014}. In
 these cases, the $2$-clique and $2$-club numbers are the same, except for ``football'' where the
@@
 really stands out here: for ``polblogs'' with $k = 3$ and $k = 4$, Wotzlaw was unable to prove
-optimality within an hour, but we required less than a second to do so. In both of these cases the
-$k$-clique and $k$-club numbers are the same, which suggests a potential improvement to $k$-club
-algorithms: first solve the maximum $k$-clique problem instead, and test whether the result found is
-a $k$-club, before embarking upon a more complicated search. (Note that a negative result does not
-imply that the $k$-clique and $k$-club numbers necessarily differ, since solutions are not unique.)
+optimality within an hour, but we required less than two seconds to do so. In both of these cases the
+$k$-clique and $k$-club numbers are the same.
 
@@
 \end{table}
+
+Parallel search is an active research area in maximum clique algorithms
+\cite{DBLP:journals/algorithms/McCreeshP13,DBLP:journals/jcisd/DepolliKRTJ13,DBLP:journals/topc/McCreeshP15,DBLP:journals/cor/SegundoLP16}.
+The idea is to work with a shared incumbent, and to speculatively explore portions of the search
+tree in parallel using multiple threads, in the hopes that most of the speculative work will
+contribute to the optimality proof.
+
+We selected the four graphs which we were unable to solve (for some values of $k$), along with two
+of the challenging graphs which required substantial amounts of search (at least half a million
+nodes) to solve, and repeated the experiments using parallel search with 20 threads, a timeout of 12
+hours, and the domination rule (with the laziness made thread-safe) enabled. The results are shown
+in \cref{table:parallel}.  Parallel search allowed us to close one more instance (``cond-mat'' with
+$k=3$), and improved the bounds on ??  further instances. However, to get sufficient work balance,
+we had to use the resplitting strategy proposed by \citeauthor{DBLP:journals/topc/McCreeshP15}
+\shortcite{DBLP:journals/topc/McCreeshP15}, and we had
+to allow for work splitting to depth 10 rather than depth 3. Simpler static
+decomposition approaches \cite{DBLP:journals/jcisd/DepolliKRTJ13,DBLP:journals/cor/SegundoLP16}
+would give little to no speedup on many harder $k$-clique instances, unless it were somehow possible
+to generate $\mathcal{O}(|V|^{10})$ subproblems.
+
+For the easier instances where the sequential runtime is known, the parallel search did more
+work---this is to be expected, since the work distribution approach we used recomputed parts of the
+search space (to allow for more efficient mutable data structures to be used during search), and
+speculative parallelism is unlikely to contribute to the solution when the sequential search tree is
+small. However, despite the extra work, and despite having to introduce overhead into early stages
+of the algorithm to allow for work stealing, and despite having more processing power but not more
+memory bandwidth, in no cases were the parallel runtimes substantially longer than the sequential
+runtimes (although in many cases they were not substantially better either). We also did not
+parallelise the construction of $G^k$ or the preprocessing stage of the algorithm, which in many
+cases dominated the runtime.
+
+On the other hand, for ``wing-nodal'' with $k=4$, the parallel run did much less work than the
+sequential run: a speedup of over 100 was obtained from 20 cores. A similar effect occurred with
+``cond-mat'' and $k=4$, and on several of the graphs where the incumbent was improved, it was
+improved after less time than the sequential runs took. This is because the work splitting mechanism
+we used explicitly diversifies at the top of search first, where branching heuristics are least
+likely to be correct, which leads to an initial incumbent being found faster---this is in line with
+recent observations that tailored work stealing should be favoured over randomised work stealing for
+combinatorial search problems \cite{DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}.
 
@@
     \label{figure:graph-omega}
-\end{figure}
 
-\begin{figure}[t]
+    \vspace*{2em}
     \centering
@@
     \label{figure:graph-omega-2}
-\end{figure}
 
-\begin{figure}[t]
+    \vspace*{2em}
     \centering

### 2016-04-08 paper/paper.tex
Marketing

@@
 
-We introduced a new lazy global domination rule. This was sometimes extremely beneficial---without
-this rule, we would have been unable to solve six of the problem instances we considered, and many
-others would have taken much longer. However, even with laziness there is still sometimes a cost to
-pay when this rule does nothing. This rule is thus harmful for the graphs typically considered for
-the maximum clique problem, and we see the benefit of tailoring algorithms to the problem being
-solved. We suggest that a similar rule may also be useful for the maximum $k$-club problem.
+We introduced a new lazy global domination rule. This was sometimes extremely beneficial, leading to
+exponential reductions in the search space---without this rule, we would have been unable to solve
+eight of the problem instances we considered, and many others would have taken much longer. However,
+even with laziness there is still sometimes a cost to pay when this rule does nothing. This rule is
+thus harmful (although only by a polynomial factor) for the graphs typically considered for the
+maximum clique problem, and we see the benefit of tailoring algorithms to the problem being solved.
+We suggest that a similar rule may also be useful for the maximum $k$-club problem.
 

### 2016-04-11 paper/paper.tex
Tweaking

@@
 \cite{Bomze:1999,Butenko:2006}. However, when analysing real-world data, a clique may be too strong
-a requirement. A $k$-clique\footnote{Or sometimes $n$-clique or $s$-clique; in an unfortunate clash
-of notation, ``$k$-clique problem'' is sometimes used elsewhere for the decision version of the
-clique problem, to distinguish it from the maximum clique problem.} is a relaxed form of clique,
-where instead of requiring each pair of vertices to be directly adjacent, we only require they be
+a requirement. A $k$-clique (or sometimes $n$-clique or $s$-clique---in an unfortunate clash of
+notation, ``$k$-clique problem'' is sometimes used elsewhere for the decision version of the clique
+problem, to distinguish it from the maximum clique problem) is a relaxed form of clique, where
+instead of requiring each pair of vertices to be directly adjacent, we only require they be
 connected by a path of length at most $k$ \cite{Luce:1950}.  Thus a $1$-clique \emph{is} a clique, a
@@
 
-Let $v$ and $w$ be distinct vertices in a graph $G$ (they may or may not be adjacent). We say that
-$v$ \emph{dominates} $w$ if the neighbourhood of $w$, excluding $v$, is a (possibly non-strict)
-subset of the neighbourhood of $v$, excluding $w$. From a maximum clique perspective, this means
-that $v$ is ``better than'' $w$. If $v$ and $w$ are adjacent, any clique containing $w$ may always
-be extended by the inclusion of $v$; if $v$ and $w$ are non-adjacent, replacing $w$ with $v$ in any
-clique containing $w$ cannot reduce the amount by which the clique may be grown.
+Let $v$ and $w$ be distinct vertices in a graph $G$. We say that $v$ \emph{dominates} $w$ if the
+neighbourhood of $w$, excluding $v$, is a (possibly non-strict) subset of the neighbourhood of $v$,
+excluding $w$. From a maximum clique perspective, this means that $v$ is ``better than'' $w$. If $v$
+and $w$ are adjacent, any clique containing $w$ may always be extended by the inclusion of $v$; if
+$v$ and $w$ are non-adjacent, replacing $w$ with $v$ in any clique containing $w$ cannot reduce the
+amount by which the clique may be grown.
 
@@
 
-Seven of these graphs were considered for the $k$-club problem by Wotzlaw \shortcite{Wotzlaw:2014}. In
-these cases, the $2$-clique and $2$-club numbers are the same, except for ``football'' where the
+Seven of these graphs were considered for the $k$-club problem by Wotzlaw \shortcite{Wotzlaw:2014}.
+In these cases, the $2$-clique and $2$-club numbers are the same, except for ``football'' where the
 $2$-club number is 16 but the $2$-clique number is 17; for $k = 3$ and $k = 4$ there are some
-differences. The difference in computational difficulty between the $k$-clique and $k$-club problems
-really stands out here: for ``polblogs'' with $k = 3$ and $k = 4$, Wotzlaw was unable to prove
-optimality within an hour, but we required less than two seconds to do so. In both of these cases the
+differences. There is a large difference in computational difficulty between the $k$-clique and
+$k$-club problems: for ``polblogs'' with $k = 3$ and $k = 4$, Wotzlaw was unable to prove optimality
+within an hour, but we required less than two seconds to do so. In both of these cases the
 $k$-clique and $k$-club numbers are the same.
@@
 
-Parallel search is an active research area in maximum clique algorithms
+Parallel search is an active research area for maximum clique algorithms
 \cite{DBLP:journals/algorithms/McCreeshP13,DBLP:journals/jcisd/DepolliKRTJ13,DBLP:journals/topc/McCreeshP15,DBLP:journals/cor/SegundoLP16}.
@@
 of the challenging graphs which required substantial amounts of search (at least half a million
-nodes) to solve, and repeated the experiments using parallel search with 20 threads, a timeout of 12
-hours, and the domination rule (with the laziness made thread-safe) enabled. The results are shown
-in \cref{table:parallel}.  Parallel search allowed us to close one more instance (``cond-mat'' with
-$k=3$), and improved the bounds on ??  further instances. However, to get sufficient work balance,
-we had to use the resplitting strategy proposed by \citeauthor{DBLP:journals/topc/McCreeshP15}
-\shortcite{DBLP:journals/topc/McCreeshP15}, and we had
-to allow for work splitting to depth 10 rather than depth 3. Simpler static
-decomposition approaches \cite{DBLP:journals/jcisd/DepolliKRTJ13,DBLP:journals/cor/SegundoLP16}
-would give little to no speedup on many harder $k$-clique instances, unless it were somehow possible
-to generate $\mathcal{O}(|V|^{10})$ subproblems.
+nodes) to solve, and repeated the experiments using parallel search with 20 threads and the
+domination rule (with the laziness made thread-safe) enabled, using the resplitting strategy
+described by \citeauthor{DBLP:journals/topc/McCreeshP15} \shortcite{DBLP:journals/topc/McCreeshP15}.
+The results are shown in \cref{table:parallel}. Parallel search allowed us to close two more
+instances, and improved the bounds on ?? further instances with 12 hours of search. However, to get
+sufficient work balance, we had to allow for work splitting to depth 10 rather than depth 3. A close
+inspection of the search patterns showed that simpler static decomposition approaches
+\cite{DBLP:journals/jcisd/DepolliKRTJ13,DBLP:journals/cor/SegundoLP16} would give little to no
+speedup on many harder $k$-clique instances, unless it were somehow possible to generate
+$\mathcal{O}(|V|^{10})$ subproblems.
 
@@
 sequential run: a speedup of over 100 was obtained from 20 cores. A similar effect occurred with
-``cond-mat'' and $k=4$, and on several of the graphs where the incumbent was improved, it was
-improved after less time than the sequential runs took. This is because the work splitting mechanism
-we used explicitly diversifies at the top of search first, where branching heuristics are least
-likely to be correct, which leads to an initial incumbent being found faster---this is in line with
-recent observations that tailored work stealing should be favoured over randomised work stealing for
+``cond-mat'' and $k=4$, and on ?? of the graphs where the incumbent was improved, it was improved
+after less time than the sequential runs took. This is because the work splitting mechanism we used
+explicitly diversifies at the top of search first, where branching heuristics are least likely to be
+correct, which leads to an initial incumbent being found faster---this is in line with recent
+observations that tailored work stealing should be favoured over randomised work stealing for
 combinatorial search problems \cite{DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}.
@@
 \subsection{Random Graphs}
-
-An Erd\H{o}s-R\'{e}nyi random graph $G(n, p)$ has $n$ vertices, and an edge between each distinct
-pair of vertices with probability $p$, chosen independently. Here we investigate the size of a
-maximum $k$-clique in such graphs, and the complexity of finding it. In each case, we use an average
-over 100 samples for every point.  We do not use the domination rule for these experiments: the
-probability of random graphs having dominating vertices is very low.
-
-In \cref{figure:graph-omega} we illustrate the average value of $\tilde{\omega}_k$ in $G(200,
-p)$ for different values of $k$, and a range of values of $p$ for the $x$-axis. We see that even for
-very low edge probabilities, a maximum $k$-clique quickly covers the entire graph.  (This is in contrast to
-the maximum clique problem, where a maximum clique does not even cover a quarter of the graph for
-edge probabilities below 0.75.) In \cref{figure:graph-nodes} we show the average size of the
-search space (number of nodes, or recursive calls made) for the same problem. We see that there is a
-complexity peak for each $k$, although the peak is much smaller for $k = 4$ than it is for $k = 3$,
-which is in turn much smaller than it is for $k = 2$. The peak also occurs for lower edge
-probabilities as $k$ increases. For contrast, for the maximum clique problem, the peak occurs at
-around edge probability 0.9, and is two orders of magnitude larger.
 
@@
 \end{figure}
+
+An Erd\H{o}s-R\'{e}nyi random graph $G(n, p)$ has $n$ vertices, and an edge between each distinct
+pair of vertices with probability $p$, chosen independently. Here we investigate the size of a
+maximum $k$-clique in such graphs, and the complexity of finding it. In each case, we use an average
+over 100 samples for every point.  We do not use the domination rule for these experiments: the
+probability of random graphs having dominating vertices is very low.
+
+In \cref{figure:graph-omega} we illustrate the average value of $\tilde{\omega}_k$ in $G(200,
+p)$ for different values of $k$, and a range of values of $p$ for the $x$-axis. We see that even for
+very low edge probabilities, a maximum $k$-clique quickly covers the entire graph.  (This is in contrast to
+the maximum clique problem, where a maximum clique does not even cover a quarter of the graph for
+edge probabilities below 0.75.) In \cref{figure:graph-nodes} we show the average size of the
+search space (number of nodes, or recursive calls made) for the same problem. We see that there is a
+complexity peak for each $k$, although the peak is much smaller for $k = 4$ than it is for $k = 3$,
+which is in turn much smaller than it is for $k = 2$. The peak also occurs for lower edge
+probabilities as $k$ increases. For contrast, for the maximum clique problem, the peak occurs at
+around edge probability 0.9, and is two orders of magnitude larger.
 
@@
 graph $G$ by considering $G^k$ in place of $G$ is feasible in practice. This is despite $G^k$
-potentially being dense even if $G$ is sparse.
+potentially being dense even if $G$ is sparse---this ruled out the use of maximum clique algorithms
+which are designed for sparse graphs
+\cite{DBLP:conf/waw/PattabiramanPGLC13,DBLP:journals/cor/SegundoLP16}, and we were considering
+graphs with many more vertices than is typical for dense maximum clique algorithms.
 
@@
 We suggest that a similar rule may also be useful for the maximum $k$-club problem.
+
+We were able to use parallelism to close two further instances, although this required a much finer
+level of task granularity than usual.  We suspect further progress could be made, for example, by
+tailoring the initial vertex ordering based upon what we know about the graphs, or by increasing the
+number of recursive calls per second by making the colouring stage cheaper, for example by reusing
+colour classes \cite{DBLP:conf/lion/NikolaevBS15}.
 

### 2016-04-12 paper/paper.tex
Missing words

@@
 The results are shown in \cref{table:parallel}. Parallel search allowed us to close two more
-instances, and improved the bounds on ?? further instances with 12 hours of search. However, to get
-sufficient work balance, we had to allow for work splitting to depth 10 rather than depth 3. A close
-inspection of the search patterns showed that simpler static decomposition approaches
+instances, and gave improved bounds on the four remaining instances within 12 hours. However, to
+get sufficient work balance, we had to allow for work splitting to depth 10 rather than depth 3. A
+close inspection of the search patterns showed that simpler static decomposition approaches
 \cite{DBLP:journals/jcisd/DepolliKRTJ13,DBLP:journals/cor/SegundoLP16} would give little to no
@@
 sequential run: a speedup of over 100 was obtained from 20 cores. A similar effect occurred with
-``cond-mat'' and $k=4$, and on ?? of the graphs where the incumbent was improved, it was improved
-after less time than the sequential runs took. This is because the work splitting mechanism we used
-explicitly diversifies at the top of search first, where branching heuristics are least likely to be
-correct, which leads to an initial incumbent being found faster---this is in line with recent
-observations that tailored work stealing should be favoured over randomised work stealing for
-combinatorial search problems \cite{DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}.
+``cond-mat'' and $k=4$. This is because the work splitting mechanism we used explicitly diversifies
+at the top of search first, where branching heuristics are least likely to be correct, which leads
+to an initial incumbent being found faster---this is in line with recent observations that tailored
+work stealing should be favoured over randomised work stealing for combinatorial search problems
+\cite{DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}.
 

### 2016-04-12 paper/paper.tex
Little wording changes

@@
 maximum clique perspective these would be considered unusually large graphs. However, only five were
-unsolvable within an hour (plus a sixth when the domination rule was not used), and over half of the
+unsolvable within an hour (plus a further two when the domination rule was not used), and over half of the
 problems took under two seconds.
@@
 
-On the other hand, for ``wing-nodal'' with $k=4$, the parallel run did much less work than the
-sequential run: a speedup of over 100 was obtained from 20 cores. A similar effect occurred with
-``cond-mat'' and $k=4$. This is because the work splitting mechanism we used explicitly diversifies
-at the top of search first, where branching heuristics are least likely to be correct, which leads
-to an initial incumbent being found faster---this is in line with recent observations that tailored
-work stealing should be favoured over randomised work stealing for combinatorial search problems
+As well as improving the results on the instances we could not solve sequentially, parallel search
+sometimes gave large improvements for the easier instances. For ``wing-nodal'' with $k=4$, the
+parallel run did much less work than the sequential run: a speedup of over 100 was obtained from 20
+cores. A similar effect occurred with ``cond-mat'' and $k=4$. This is because the work splitting
+mechanism we used explicitly diversifies at the top of search first, where branching heuristics are
+least likely to be correct, which leads to an initial incumbent being found faster---this is in line
+with recent observations that tailored work stealing should be favoured over randomised work
+stealing for combinatorial search problems
 \cite{DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}.
@@
 An Erd\H{o}s-R\'{e}nyi random graph $G(n, p)$ has $n$ vertices, and an edge between each distinct
-pair of vertices with probability $p$, chosen independently. Here we investigate the size of a
+pair of vertices with probability $p$, chosen independently. We now investigate the size of a
 maximum $k$-clique in such graphs, and the complexity of finding it. In each case, we use an average
@@
 which are designed for sparse graphs
-\cite{DBLP:conf/waw/PattabiramanPGLC13,DBLP:journals/cor/SegundoLP16}, and we were considering
+\cite{DBLP:conf/waw/PattabiramanPGLC13,DBLP:journals/cor/SegundoLP16}, and we were working with
 graphs with many more vertices than is typical for dense maximum clique algorithms.
@@
 exponential reductions in the search space---without this rule, we would have been unable to solve
-eight of the problem instances we considered, and many others would have taken much longer. However,
+nine of the problem instances we considered, and many others would have taken much longer. However,
 even with laziness there is still sometimes a cost to pay when this rule does nothing. This rule is
@@
 We were able to use parallelism to close two further instances, although this required a much finer
-level of task granularity than usual.  We suspect further progress could be made, for example, by
-tailoring the initial vertex ordering based upon what we know about the graphs, or by increasing the
-number of recursive calls per second by making the colouring stage cheaper, for example by reusing
-colour classes \cite{DBLP:conf/lion/NikolaevBS15}.
+level of task granularity than usual.  We suspect further progress could be made by tailoring the
+initial vertex ordering based upon what we know about the graphs, or by increasing the number of
+recursive calls per second by making the colouring stage cheaper, for example by reusing colour
+classes \cite{DBLP:conf/lion/NikolaevBS15}.
 

### 2016-05-17 paper/paper.tex
Deanonymous

@@
 \setlength{\pdfpageheight}{11in}
-% \pdfinfo{
-% /Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
-% /Author (Ciaran McCreesh, Patrick Prosser)}
 \pdfinfo{
 /Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
-/Author (First Author and Second Author)}
+/Author (Ciaran McCreesh, Patrick Prosser)}
 \setcounter{secnumdepth}{0}
@@
 \title{Finding Maximum $k$-Cliques Faster using Lazy Global Domination}
-% \author{
-%    Ciaran McCreesh%
-%    \thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/K503058/1]}
-%    \and
-%    Patrick Prosser
-%    \\ University of Glasgow, Glasgow, Scotland
-%}
  \author{
-    First Author%
-    \thanks{This work was supported by a pile of unmarked bills from an anonymous funding agency}
+    Ciaran McCreesh%
+    \thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/K503058/1]}
     \and
-    Second Author
-    \\ Secret Department of Secrecy, Some Institution, Somewhere
+    Patrick Prosser
+    \\ University of Glasgow, Glasgow, Scotland
 }

### 2016-05-17 paper/paper.tex
Larger fonts

@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{6pt} % carefully tuned!
+    \scriptsize\setlength{\tabcolsep}{5pt} % carefully tuned!
     \centering
@@
 \caption{Experimental results for DIMACS clique graphs with diameter greater
-    than two. For each graph, we consider $k$ equal to 2, 3 and 4. In each case
-    we show the density of $G^k$, the size of a maximum $k$-clique, and then
-    for both the unmodified algorithm and the algorithm with our lazy global
-domination step, the number of nodes required, and the runtime in seconds.}\label{table:clique}
+    than two.}\label{table:clique}
 \end{table}
@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{4pt} % carefully tuned!
+    \scriptsize\setlength{\tabcolsep}{3.5pt} % carefully tuned!
+    \setlength{\aboverulesep}{-0.4pt} % this only affects the cmidrules on this table
     \centering
@@
         & & & & \multicolumn{2}{c}{Unmodified} & \multicolumn{2}{c}{With Domination} \\
-    \cmidrule(lr){5-6}
-    \cmidrule(lr){7-8}
+        \cmidrule(lr){5-6} \cmidrule(lr){7-8}
     Instance & \multicolumn{1}{c}{$k$} & \multicolumn{1}{c}{$D$} & \multicolumn{1}{c}{$\tilde{\omega}_k$} &
@@
 \end{tabular}
-\caption{Experimental results for the smaller DIMACS partitioning graphs. For
-    each graph, we consider $k$ equal to 2, 3 and 4. In each case we show the
-    density of $G^k$, the size of a maximum $k$-clique, and then for both the
-    unmodified algorithm and the algorithm with our lazy global domination
-step, the number of nodes required, and the runtime in seconds.}\label{table:partitioning}
+    \caption{Results for smaller DIMACS partitioning graphs.}\label{table:partitioning}
 \end{table}
@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{5pt} % carefully tuned!
+    \scriptsize\setlength{\tabcolsep}{3.8pt} % carefully tuned!
+    \setlength{\aboverulesep}{-0.4pt} % this only affects the cmidrules on this table
     \centering
@@
 \end{tabular}
-\caption{Experimental results for the smaller DIMACS clustering graphs. For
-    each graph, we consider $k$ equal to 2, 3 and 4. In each case we show the
-    density of $G^k$, the size of a maximum $k$-clique, and then for both the
-    unmodified algorithm and the algorithm with our lazy global domination
-step, the number of nodes required, and the runtime in seconds.}\label{table:clustering}
+\caption{Results for smaller DIMACS clustering graphs.}\label{table:clustering}
 \end{table}
@@
 \begin{table}
-    \tiny\setlength{\tabcolsep}{4.6pt} % carefully tuned!
+    \scriptsize\setlength{\tabcolsep}{3.4pt} % carefully tuned!
     \centering

### 2016-05-17 paper/paper.tex
No manual figure spacing
AAAI style caps Using

@@
 \pdfinfo{
-/Title (Finding Maximum $k$-Cliques Faster using Lazy Global Domination)
+/Title (Finding Maximum $k$-Cliques Faster Using Lazy Global Domination)
 /Author (Ciaran McCreesh, Patrick Prosser)}
@@
 \begin{document}
-\title{Finding Maximum $k$-Cliques Faster using Lazy Global Domination}
+\title{Finding Maximum $k$-Cliques Faster Using Lazy Global Domination}
  \author{
@@
     \label{figure:graph-omega}
+\end{figure}
 
-    \vspace*{2em}
+\begin{figure}[t]
     \centering
@@
     \label{figure:graph-omega-2}
+\end{figure}
 
-    \vspace*{2em}
+\begin{figure}[t]
     \centering

### 2016-05-17 paper/paper.tex
Address review 2

@@
 problem, to distinguish it from the maximum clique problem) is a relaxed form of clique, where
-instead of requiring each pair of vertices to be directly adjacent, we only require they be
+instead of requiring each pair of vertices to be directly adjacent, we only require that they be
 connected by a path of length at most $k$ \cite{Luce:1950}.  Thus a $1$-clique \emph{is} a clique, a
@@
 A maximum clique algorithm can easily be adapted to find a maximum $k$-clique in a graph $G$ by
-considering the graph $G^k$, which we describe below. However, it is not obvious that this is a
-viable approach: even if $G$ is sparse, $G^k$ may not be, and the maximum clique problem on dense
-graphs can be very challenging computationally. Here we take a state-of-the-art maximum clique
-algorithm which is suitable for use on dense graphs \cite{Prosser:2012}, and
-investigate whether this approach is feasible in practice. We modify the algorithm to include a new
-lazy ``global domination'' inference step---this technique provides no benefit for typical maximum
-clique problems, but for maximum $k$-clique graphs it sometimes gives improvements of several orders
-of magnitude. We present computational results for the maximum $k$-clique problem on a range of
-benchmark and real-world graphs. We finish with a detailed look at random graphs.
+considering the graph $G^k$: this graph has the same vertex set as $G$, and edges between any two
+distinct vertices $v_1$ and $v_2$ iff there is a path of length at most $k$ between $v_1$ and $v_2$
+in $G$. It is easy to see that maximum cliques in $G^k$ correspond with maximum $k$-cliques in $G$
+\cite{Balasundaram:2005}.  However, it is not obvious that this is a viable approach: even if $G$ is
+sparse, $G^k$ may not be, and the maximum clique problem on dense graphs can be very challenging
+computationally. Here we take a state-of-the-art maximum clique algorithm which is suitable for use
+on dense graphs \cite{Prosser:2012}, and investigate whether this approach is feasible in practice.
+We modify the algorithm to include a new lazy ``global domination'' inference step---this technique
+provides no benefit for typical maximum clique problems, but for maximum $k$-clique graphs it
+sometimes gives improvements of several orders of magnitude. We present computational results for
+the maximum $k$-clique problem on a range of benchmark and real-world graphs. We finish with a
+detailed look at random graphs.
 
@@
 Our approach for finding a maximum $k$-clique is presented as \cref{algorithm:maxKClique}. Our first
-step (\mcline{power}) is to replace our input graph $G$ with a modified graph $G^k$. This graph has
-the same vertex set as $G$, and edges between any two distinct vertices $v_1$ and $v_2$ iff there is
-a path of length at most $k$ between $v_1$ and $v_2$ in $G$. We may construct this graph using a
-bounded breadth-first search: we refer to \citeauthor{Chang:2013} \shortcite{Chang:2013} for how to
-implement this quickly in practice. Now it is easy to see that maximum cliques in $G^k$ correspond
-with maximum $k$-cliques in $G$ \cite{Balasundaram:2005}.
+step (\mcline{power}) is to replace our input graph $G$ with $G^k$. We may construct this graph
+using a bounded breadth-first search: we refer to \citeauthor{Chang:2013} \shortcite{Chang:2013} for
+how to implement this quickly in practice.
 
@@
 
-By comparing these results with the $k$-club results of Chang et al.\ \shortcite{Chang:2013}, we see that
-in all but four cases the $k$-clique and $k$-club numbers are equal; all of these differences occur
-when $k = 4$. (Chang et al.\ did not investigate the ``Erdos02'' graph, but Wotzlaw
-\shortcite{Wotzlaw:2014} confirmed privately that the $k$-clique and $k$-club numbers are the same here
-too.) On the other hand, the $k$-clique numbers are sometimes much easier to find, both
+By comparing these results with the $k$-club results of Chang et al.\ \shortcite{Chang:2013}, we see
+that in all but four cases the $k$-clique and $k$-club numbers are equal; all of these differences
+occur when $k = 4$. (Chang et al.\ did not investigate the ``Erdos02'' graph, but
+\citeauthor{Wotzlaw:2014} confirmed privately that the $k$-clique and $k$-club numbers are the same
+here too.) On the other hand, the $k$-clique numbers are sometimes much easier to find, both
 algorithmically and computationally.

### 2016-05-17 paper/paper.tex
Less wordy

@@
 
-The key step in \citeauthor{Tomita:2010}'s algorithms is to produce a constructive colouring, which
-is used in a clever way.  The $\colourOrder$ routine does not just return the number of colours
-used. Instead, it returns a pair of arrays, $\order$ and $\bounds$. The $\order$ array contains
-vertices, in the order in which they were coloured. The $i$th entry of the $\bounds$ array contains
-the colour number used for the $i$th vertex in $\order$. We illustrate this in
-\cref{figure:colours}. Crucially, $\bounds$ is non-decreasing (i.e.\ $\bounds[i+1] \ge \bounds[i]$),
-and we may colour the subgraph induced by the first $i$ vertices of $\order$ using $\bounds[i]$
-colours.
+The key step in \citeauthor{Tomita:2010}'s algorithms is to produce a constructive colouring---the
+$\colourOrder$ routine does not just return the number of colours used. Instead, it returns a pair
+of arrays, $\order$ and $\bounds$. The $\order$ array contains vertices, in the order in which they
+were coloured. The $i$th entry of the $\bounds$ array contains the colour number used for the $i$th
+vertex in $\order$. We illustrate this in \cref{figure:colours}. Crucially, $\bounds$ is
+non-decreasing (i.e.\ $\bounds[i+1] \ge \bounds[i]$), and we may colour the subgraph induced by the
+first $i$ vertices of $\order$ using $\bounds[i]$ colours.
 

### 2016-05-17 paper/paper.tex
Review 3 comments

@@
 % cref style
-\crefname{algocf}{algorithm}{algorithm}
+\crefname{algocf}{Algorithm}{Algorithm}
 \Crefname{algocf}{Algorithm}{Algorithms}
-\crefname{figure}{figure}{figures}
+\crefname{figure}{Figure}{Figures}
 \Crefname{figure}{Figure}{Figures}
-\crefname{table}{table}{tables}
+\crefname{table}{Table}{Tables}
 \Crefname{table}{Table}{Tables}
@@
 A recent survey by \citeauthor{Shahinpour:2013} \shortcite{Shahinpour:2013} discusses algorithms and
-results for $k$-clique and $k$-club problems .  We adopt their notation of $\tilde{\omega}_k$ for
+results for $k$-clique and $k$-club problems.  We adopt their notation of $\tilde{\omega}_k$ for
 the size of a maximum $k$-clique; the use of $\omega$ for the size of a maximum clique is standard.
@@
 We now give experimental results on a range of standard benchmarks, and on real-world and random
-graphs. Experiments were run on a machine with dual Intel E5-2687W v3 processors and 256GBytes
-RAM; single-threaded runtimes are given, except in \cref{table:parallel}, where 20 threads were used
-(our machine does not have hyper-threading enabled). Our software was implemented in C++, using
-C++11 native threads, and was compiled using GCC 5.1.0. The time taken to read in the graph from a
-file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
+graphs. Experiments were run on a machine with dual 3.1GHz Intel E5-2687W v3 processors and
+256GBytes RAM; single-threaded runtimes are given, except in \cref{table:parallel}, where 20 threads
+were used (our machine does not have hyper-threading enabled). Our software was implemented in C++,
+using C++11 native threads, and was compiled using GCC 5.1.0. The time taken to read in the graph
+from a file is excluded, but preprocessing time (including the construction of $G^k$ and the bitset
 encoding) is included. We use the term \emph{nodes} to refer to the number of recursive calls made

### 2016-05-23 paper/paper.tex
AAAIify

@@
 
-\usepackage{cleveref}
-\usepackage{complexity}
+\usepackage{cleveref}                    % \cref
+\usepackage{complexity}                  % \NP
 \usepackage[ruled,vlined]{algorithm2e}   % algorithms
 \usepackage{booktabs}                    % less ugly tables
-\usepackage[tight-spacing=true]{siunitx}
+\usepackage[tight-spacing=true]{siunitx} % formatting of units
 \usepackage{tikz}                        % For pretty pictures
-\usepackage{gnuplot-lua-tikz}          % graphs
-
-% \usepackage{showframe}
+\usepackage{graphicx}                    % For the charts
 
@@
 decorations.pathreplacing, patterns, tikzmark, fit}
-
-% \input isn't expandable. yay.
-\makeatletter\newcommand*\inputhaxx[1]{\@@input #1}\makeatother
 
@@
 \tikzset{selectedvertex/.style={vertex, fill=safedarkblue}}
-\tikzset{delvertex/.style={vertex, dotted, color=safelightgrey}}
 \tikzset{vertexc1/.style={vertex, fill=safelightblue}}
@@
 \tikzset{vertexc4/.style={vertex, fill=safedarkorange, text=safenearlywhite}}
-
 \tikzset{edge/.style={color=safelightgrey}}
 \tikzset{bedge/.style={very thick}}
-\tikzset{deledge/.style={dotted, color=safelightgrey}}
-\tikzset{edgel1/.style={color=safeverydarkblue}}
-\tikzset{edgel2/.style={color=safemediumorange}}
-\tikzset{edgel3/.style={very thick, color=safedarkblue}}
-\tikzset{edgel4/.style={very thick, color=safelightorange}}
-
 \tikzset{processarrow/.style={->, very thick, decorate, decoration={snake, post length=0.5mm}}}
@@
 
-\makeatletter\@ifpackageloaded{showframe}{
-    % show lines down the middle columns too
-    % http://tex.stackexchange.com/questions/16199/test-if-a-package-or-package-option-is-loaded
-\newlength\Fcolumnseprule\setlength\Fcolumnseprule{0.4pt}\def\@outputdblcol{%
-  \if@firstcolumn\global \@firstcolumnfalse \global \setbox\@leftcolumn \box\@outputbox
-  \else
-    \global \@firstcolumntrue
-    \setbox\@outputbox \vbox{\hb@xt@\textwidth{\hb@xt@\columnwidth{\box\@leftcolumn \hss}%
-    \vrule \@width\Fcolumnseprule\hfil{\normalcolor\vrule \@width\columnseprule}
-    \hfil\vrule \@width\Fcolumnseprule\hb@xt@\columnwidth {\box\@outputbox \hss}}}\@combinedblfloats\@outputpage
-    \begingroup\@dblfloatplacement\@startdblcolumn\@whilesw\if@fcolmade \fi{\@outputpage\@startdblcolumn}\endgroup
-  \fi
-}}{}
-\makeatother
-
-\renewcommand{\topfraction}{0.9}
-\renewcommand{\bottomfraction}{0.8}
-\renewcommand{\dbltopfraction}{0.9}
-\renewcommand{\textfraction}{0.05}
-\setcounter{dbltopnumber}{2}
-\renewcommand{\floatpagefraction}{0.85}
-\renewcommand{\dblfloatpagefraction}{0.85}
-
-% \frenchspacing
 \setlength{\pdfpagewidth}{8.5in}
@@
 
-\begin{algorithm}[t]\DontPrintSemicolon
+\begin{algorithm}[t!]\DontPrintSemicolon
     \begin{tikzpicture}[remember picture,overlay]
@@
     \midrule
-    \inputhaxx{gen-table-erdos}
+Erdos971  & 2&0.09&42&42&0.0&42&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}472$& 3&0.31&117&121&0.0&119&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}1314$& 4&0.56&235&468&0.0&468&0.0\\
+Erdos972  & 2&0.01&258&258&0.5&258&0.5\\
+\hspace*{0.2em}\color{gray}$|V|{=}5488$& 3&0.09&517&537&0.8&521&0.8\\
+\hspace*{0.2em}\color{gray}$|E|{=}8972$& 4&0.35&1509&\color{gray}\num{1.1e+07}&\color{gray}$>$ 1 h&8197&3.8\\
+Erdos981  & 2&0.09&43&43&0.0&43&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}485$& 3&0.31&123&358&0.0&354&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}1381$& 4&0.57&245&246&0.0&246&0.0\\
+Erdos982  & 2&0.01&274&274&0.6&274&0.6\\
+\hspace*{0.2em}\color{gray}$|V|{=}5822$& 3&0.09&547&555&0.9&547&0.9\\
+\hspace*{0.2em}\color{gray}$|E|{=}9505$& 4&0.35&1594&\color{gray}\num{1.1e+07}&\color{gray}$>$ 1 h&618826&207.3\\
+Erdos991  & 2&0.09&43&44&0.0&44&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}492$& 3&0.31&126&375&0.0&374&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}1417$& 4&0.57&246&491&0.0&491&0.0\\
+Erdos992  & 2&0.01&277&277&0.6&277&0.6\\
+\hspace*{0.2em}\color{gray}$|V|{=}6100$& 3&0.09&562&573&0.9&562&0.8\\
+\hspace*{0.2em}\color{gray}$|E|{=}9939$& 4&0.35&1643&\color{gray}\num{1.0e+07}&\color{gray}$>$ 1 h&202543&69.4\\
+Erdos02  & 2&0.02&508&508&0.8&508&0.8\\
+\hspace*{0.2em}\color{gray}$|V|{=}6927$& 3&0.20&1014&1022&1.3&1015&1.3\\
+\hspace*{0.2em}\color{gray}$|E|{=}11850$& 4&1.00&6927&6927&11.9&6927&10.9\\
     \bottomrule
@@
     \midrule
-    \inputhaxx{gen-table-dimacs}
+c-fat200-1  & 2&0.13&18&41&0.0&35&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}200$& 3&0.19&24&74&0.0&48&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}1534$& 4&0.24&30&134&0.0&65&0.0\\
+c-fat200-2  & 2&0.27&35&35&0.0&35&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}200$& 3&0.39&46&488&0.0&102&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}3235$& 4&0.50&57&1496&0.0&128&0.0\\
+c-fat200-5  & 2&0.71&87&11513&0.0&257&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}200$& 3&1.00&200&200&0.0&200&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}8473$& 4&1.00&200&200&0.0&200&0.0\\
+c-fat500-1  & 2&0.06&21&52&0.0&43&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}500$& 3&0.09&28&28&0.0&28&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}4459$& 4&0.11&35&35&0.0&35&0.0\\
+c-fat500-2  & 2&0.12&39&134&0.0&79&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}500$& 3&0.17&52&52&0.0&52&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}9139$& 4&0.22&65&65&0.0&65&0.0\\
+c-fat500-5  & 2&0.31&96&10133&0.1&196&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}500$& 3&0.44&128&128&0.1&128&0.1\\
+\hspace*{0.2em}\color{gray}$|E|{=}23191$& 4&0.56&159&\color{gray}\num{1.5e+09}&\color{gray}$>$ 1 h&326&0.1\\
+c-fat500-10  & 2&0.62&189&\color{gray}\num{1.1e+09}&\color{gray}$>$ 1 h&560&0.1\\
+\hspace*{0.2em}\color{gray}$|V|{=}500$& 3&0.87&252&252&0.1&252&0.1\\
+\hspace*{0.2em}\color{gray}$|E|{=}46627$& 4&1.00&500&500&0.1&500&0.1\\
+p-hat300-1  & 2&1.00&299&299&0.0&299&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}300$& 3&1.00&300&300&0.1&300&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}10933$& 4&1.00&300&300&0.0&300&0.0\\
     \bottomrule
@@
     \scriptsize\setlength{\tabcolsep}{3.5pt} % carefully tuned!
-    \setlength{\aboverulesep}{-0.4pt} % this only affects the cmidrules on this table
+    % The following only affects the cmidrules on this table: by default booktabs sticks larger
+    % spacing than normal latex tables do above and below mid-table rules. Usually that's fine, but
+    % here the table is just slightly too large for the page, and I'd rather not use a smaller font
+    % or cut one of the data points. (I suppose the other option would just be to just cut the
+    % cmidrules out entirely, and then it fits perfectly...)
+    \setlength{\aboverulesep}{-0.4pt}
     \centering
@@
     \midrule
-    \inputhaxx{gen-table-dimacs10walshaw}
+3elt  & 2&0.00&10&340&0.3&340&0.6\\
+\hspace*{0.2em}\color{gray}$|V|{=}4720$& 3&0.01&16&1582&0.4&1582&1.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}13722$& 4&0.01&27&911&0.4&911&0.7\\
+4elt  & 2&0.00&11&486&3.8&486&5.5\\
+\hspace*{0.2em}\color{gray}$|V|{=}15606$& 3&0.00&20&717&3.9&717&6.1\\
+\hspace*{0.2em}\color{gray}$|E|{=}45878$& 4&0.00&36&345&3.9&345&4.4\\
+add20  & 2&0.04&124&124&0.1&124&0.2\\
+\hspace*{0.2em}\color{gray}$|V|{=}2395$& 3&0.25&671&671&0.2&671&0.2\\
+\hspace*{0.2em}\color{gray}$|E|{=}7462$& 4&0.67&1454&1454&0.5&1454&0.5\\
+add32  & 2&0.00&32&32&0.4&32&0.4\\
+\hspace*{0.2em}\color{gray}$|V|{=}4960$& 3&0.01&99&286&0.4&194&0.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}9462$& 4&0.03&268&268&0.4&268&0.5\\
+bcsstk29  & 2&0.01&72&9752&3.7&963&8.7\\
+\hspace*{0.2em}\color{gray}$|V|{=}13992$& 3&0.02&132&\num{2.7e+07}&1576.2&7781&11.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}302748$& 4&0.04&210&\color{gray}\num{3.2e+07}&\color{gray}$>$ 1 h&21689&16.9\\
+bcsstk30  & 2&0.01&219&224&14.5&219&14.5\\
+\hspace*{0.2em}\color{gray}$|V|{=}28924$& 3&0.03&496&509&17.8&496&17.6\\
+\hspace*{0.2em}\color{gray}$|E|{=}1007284$& 4&0.05&843&854&23.7&845&23.3\\
+bcsstk31  & 2&0.00&189&189&20.2&189&20.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}35588$& 3&0.01&278&605&21.8&369&22.8\\
+\hspace*{0.2em}\color{gray}$|E|{=}572914$& 4&0.02&428&119640&97.9&6588&29.8\\
+bcsstk33  & 2&0.03&141&141&1.3&141&1.3\\
+\hspace*{0.2em}\color{gray}$|V|{=}8738$& 3&0.08&228&26033&6.3&1744&3.8\\
+\hspace*{0.2em}\color{gray}$|E|{=}291583$& 4&0.15&435&\num{2.0e+06}&380.4&52779&17.6\\
+crack  & 2&0.00&10&2894&1.5&2894&7.8\\
+\hspace*{0.2em}\color{gray}$|V|{=}10240$& 3&0.00&17&4996&1.7&4987&10.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}30380$& 4&0.01&31&2173&1.7&2173&7.0\\
+cs4  & 2&0.00&6&5780&8.2&5780&73.4\\
+\hspace*{0.2em}\color{gray}$|V|{=}22499$& 3&0.00&12&7812&8.4&7812&78.5\\
+\hspace*{0.2em}\color{gray}$|E|{=}43858$& 4&0.00&18&29032&9.5&29032&136.3\\
+cti  & 2&0.00&7&8918&4.5&8918&58.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}16840$& 3&0.00&15&6406&4.8&6406&45.5\\
+\hspace*{0.2em}\color{gray}$|E|{=}48232$& 4&0.01&26&62316&7.0&62316&113.6\\
+data  & 2&0.01&18&638&0.1&617&0.2\\
+\hspace*{0.2em}\color{gray}$|V|{=}2851$& 3&0.02&32&4982&0.1&4913&0.2\\
+\hspace*{0.2em}\color{gray}$|E|{=}15093$& 4&0.04&52&40095&0.4&36089&0.4\\
+fe-4elt2  & 2&0.00&13&61&1.7&61&2.1\\
+\hspace*{0.2em}\color{gray}$|V|{=}11143$& 3&0.00&20&389&1.8&389&3.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}32818$& 4&0.01&32&448&1.8&446&3.0\\
+fe-pwt  & 2&0.00&16&95&21.8&95&22.6\\
+\hspace*{0.2em}\color{gray}$|V|{=}36519$& 3&0.00&29&167&21.0&167&24.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}144794$& 4&0.00&52&224&21.1&224&23.9\\
+fe-sphere  & 2&0.00&7&14173&4.5&14173&53.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}16386$& 3&0.00&12&34328&4.9&34328&111.1\\
+\hspace*{0.2em}\color{gray}$|E|{=}49152$& 4&0.00&19&73632&6.3&73632&148.7\\
+memplus  & 2&0.02&574&574&5.8&574&5.7\\
+\hspace*{0.2em}\color{gray}$|V|{=}17758$& 3&0.26&8057&8061&49.4&8058&48.2\\
+\hspace*{0.2em}\color{gray}$|E|{=}54196$& 4&0.74&8963&8963&95.5&8963&92.4\\
+uk  & 2&0.00&5&433&0.3&433&0.7\\
+\hspace*{0.2em}\color{gray}$|V|{=}4824$& 3&0.00&8&1891&0.4&1891&0.9\\
+\hspace*{0.2em}\color{gray}$|E|{=}6837$& 4&0.01&14&2168&0.4&2168&1.0\\
+vibrobox  & 2&0.02&121&302&2.9&302&3.5\\
+\hspace*{0.2em}\color{gray}$|V|{=}12328$& 3&0.08&408&1984&8.9&1984&10.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}165250$& 4&0.26&${\ge}1094$&\color{gray}\num{7.9e+06}&\color{gray}$>$ 1 h&\color{gray}\num{7.4e+06}&\color{gray}$>$ 1 h\\
+whitaker3  & 2&0.00&9&1222&1.3&1222&4.8\\
+\hspace*{0.2em}\color{gray}$|V|{=}9800$& 3&0.00&15&3724&1.4&3724&7.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}28989$& 4&0.01&23&6530&1.5&6530&9.5\\
+wing-nodal  & 2&0.01&29&648&2.1&648&3.5\\
+\hspace*{0.2em}\color{gray}$|V|{=}10937$& 3&0.02&54&13091&4.2&13039&15.2\\
+\hspace*{0.2em}\color{gray}$|E|{=}75488$& 4&0.04&114&\num{6.0e+07}&1981.3&\num{6.0e+07}&1908.9\\
     \bottomrule
@@
     \scriptsize\setlength{\tabcolsep}{3.8pt} % carefully tuned!
-    \setlength{\aboverulesep}{-0.4pt} % this only affects the cmidrules on this table
+    % The following only affects the cmidrules on this table: by default booktabs sticks larger
+    % spacing than normal latex tables do above and below mid-table rules. Usually that's fine, but
+    % here the table is just slightly too large for the page, and I'd rather not use a smaller font
+    % or cut one of the data points. (I suppose the other option would just be to just cut the
+    % cmidrules out entirely, and then it fits perfectly...)
+    \setlength{\aboverulesep}{-0.4pt}
     \centering
@@
     \midrule
-    \inputhaxx{gen-table-dimacs10cluster}
+adjnoun  & 2&0.50&50&50&0.0&50&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}112$& 3&0.91&83&164&0.0&164&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}425$& 4&0.99&107&107&0.0&107&0.0\\
+as-22july06  & 2&0.04&2391&2391&11.7&2391&11.3\\
+\hspace*{0.2em}\color{gray}$|V|{=}22963$& 3&0.36&8455&\color{gray}374427&\color{gray}$>$ 1 h&94497&898.2\\
+\hspace*{0.2em}\color{gray}$|E|{=}48436$& 4&0.79&14911&14911&252.0&14911&250.4\\
+astro-ph  & 2&0.01&361&365&5.0&362&5.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}16706$& 3&0.10&1553&1567&15.3&1560&12.5\\
+\hspace*{0.2em}\color{gray}$|E|{=}121251$& 4&0.35&${\ge}4040$&\color{gray}\num{1.0e+06}&\color{gray}$>$ 1 h&\color{gray}\num{1.1e+06}&\color{gray}$>$ 1 h\\
+celegans-meta  & 2&0.44&238&238&0.0&238&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}453$& 3&0.89&371&371&0.0&371&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}2025$& 4&0.98&432&432&0.0&432&0.0\\
+celegansneural  & 2&0.55&135&135&0.0&135&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}297$& 3&0.95&245&245&0.0&245&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}2148$& 4&1.00&295&295&0.0&295&0.0\\
+cond-mat  & 2&0.00&108&108&4.5&108&4.5\\
+\hspace*{0.2em}\color{gray}$|V|{=}16726$& 3&0.01&250&1403&5.1&844&5.6\\
+\hspace*{0.2em}\color{gray}$|E|{=}47594$& 4&0.05&720&\color{gray}\num{7.3e+06}&\color{gray}$>$ 1 h&674453&280.6\\
+cond-mat-2003  & 2&0.00&203&204&15.7&204&16.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}31163$& 3&0.02&${\ge}629$&\color{gray}\num{6.3e+06}&\color{gray}$>$ 1 h&\color{gray}\num{6.1e+06}&\color{gray}$>$ 1 h\\
+\hspace*{0.2em}\color{gray}$|E|{=}120029$& 4&0.12&${\ge}2605$&\color{gray}\num{1.9e+06}&\color{gray}$>$ 1 h&\color{gray}\num{1.8e+06}&\color{gray}$>$ 1 h\\
+cond-mat-2005  & 2&0.00&279&279&26.8&279&26.4\\
+\hspace*{0.2em}\color{gray}$|V|{=}40421$& 3&0.03&${\ge}1060$&\color{gray}\num{2.0e+06}&\color{gray}$>$ 1 h&\color{gray}\num{1.9e+06}&\color{gray}$>$ 1 h\\
+\hspace*{0.2em}\color{gray}$|E|{=}175691$& 4&0.16&${\ge}4185$&\color{gray}562816&\color{gray}$>$ 1 h&\color{gray}585596&\color{gray}$>$ 1 h\\
+dolphins  & 2&0.32&14&14&0.0&14&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}62$& 3&0.59&30&30&0.0&30&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}159$& 4&0.77&40&40&0.0&40&0.0\\
+email  & 2&0.09&72&72&0.0&72&0.1\\
+\hspace*{0.2em}\color{gray}$|V|{=}1133$& 3&0.45&233&19031&0.3&19031&0.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}5451$& 4&0.86&654&6854&0.3&6676&0.4\\
+football  & 2&0.45&17&147&0.0&145&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}115$& 3&0.95&69&70&0.0&70&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}613$& 4&1.00&115&115&0.0&115&0.0\\
+hep-th  & 2&0.00&51&51&1.0&51&1.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}8361$& 3&0.01&125&239&1.2&176&1.3\\
+\hspace*{0.2em}\color{gray}$|E|{=}15751$& 4&0.04&347&158164&21.6&23714&4.7\\
+jazz  & 2&0.69&103&107&0.0&107&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}198$& 3&0.95&174&174&0.0&174&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}2742$& 4&0.99&192&192&0.0&192&0.0\\
+karate  & 2&0.61&18&18&0.0&18&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}34$& 3&0.86&25&25&0.0&25&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}78$& 4&0.99&33&33&0.0&33&0.0\\
+lesmis  & 2&0.43&37&37&0.0&37&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}77$& 3&0.85&58&58&0.0&58&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}254$& 4&0.99&75&75&0.0&75&0.0\\
+netscience  & 2&0.01&35&35&0.0&35&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}1589$& 3&0.01&54&54&0.0&54&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}2742$& 4&0.02&85&85&0.0&85&0.0\\
+PGPgiantcompo  & 2&0.00&206&206&1.7&206&1.7\\
+\hspace*{0.2em}\color{gray}$|V|{=}10680$& 3&0.02&423&843&2.6&841&2.7\\
+\hspace*{0.2em}\color{gray}$|E|{=}24316$& 4&0.07&1161&1161&5.9&1161&6.0\\
+polblogs  & 2&0.27&352&352&0.1&352&0.1\\
+\hspace*{0.2em}\color{gray}$|V|{=}1490$& 3&0.58&776&2210&1.2&2177&1.2\\
+\hspace*{0.2em}\color{gray}$|E|{=}16715$& 4&0.66&1127&1537&2.0&1166&1.9\\
+polbooks  & 2&0.37&28&28&0.0&28&0.0\\
+\hspace*{0.2em}\color{gray}$|V|{=}105$& 3&0.64&54&54&0.0&54&0.0\\
+\hspace*{0.2em}\color{gray}$|E|{=}441$& 4&0.86&68&68&0.0&68&0.0\\
+power  & 2&0.00&20&20&0.4&20&0.4\\
+\hspace*{0.2em}\color{gray}$|V|{=}4941$& 3&0.00&30&30&0.4&30&0.4\\
+\hspace*{0.2em}\color{gray}$|E|{=}6594$& 4&0.01&61&61&0.4&61&0.4\\
     \bottomrule
@@
     \midrule
-    \inputhaxx{gen-table-parallel}
+astro-ph  & 2&0.01&361&362&5.0 &17227&4.9 \\
+\hspace*{0.2em}\color{gray}$|V|{=}16706$& 3&0.10&1553&1560&12.5 &32614&9.5 \\
+\hspace*{0.2em}\color{gray}$|E|{=}121251$& 4&0.35&${\ge}4125$&\color{gray}\num{1.1e+06}&\color{gray}$>$ 1 h&\color{gray}\num{3.1e+08}&\color{gray}$>$ 12 h\\
+cond-mat  & 2&0.00&108&108&4.5 &16833&4.5 \\
+\hspace*{0.2em}\color{gray}$|V|{=}16726$& 3&0.01&250&844&5.6 &17933&5.4 \\
+\hspace*{0.2em}\color{gray}$|E|{=}47594$& 4&0.05&720&674453&280.6 &22604&8.7 \\
+cond-mat-2003  & 2&0.00&203&204&16.0 &31424&14.8 \\
+\hspace*{0.2em}\color{gray}$|V|{=}31163$& 3&0.02&634&\color{gray}\num{6.1e+06}&\color{gray}$>$ 1 h&\num{3.4e+08}&8131.3 \\
+\hspace*{0.2em}\color{gray}$|E|{=}120029$& 4&0.12&${\ge}2606$&\color{gray}\num{1.8e+06}&\color{gray}$>$ 1 h&\color{gray}\num{5.1e+08}&\color{gray}$>$ 12 h\\
+cond-mat-2005  & 2&0.00&279&279&26.4 &40701&24.2 \\
+\hspace*{0.2em}\color{gray}$|V|{=}40421$& 3&0.03&1060&\color{gray}\num{1.9e+06}&\color{gray}$>$ 1 h&\num{1.0e+09}&69836.7 \\
+\hspace*{0.2em}\color{gray}$|E|{=}175691$& 4&0.16&${\ge}4271$&\color{gray}585596&\color{gray}$>$ 1 h&\color{gray}\num{1.2e+08}&\color{gray}$>$ 12 h\\
+vibrobox  & 2&0.02&121&302&3.5 &12699&3.4 \\
+\hspace*{0.2em}\color{gray}$|V|{=}12328$& 3&0.08&408&1984&10.0 &17246&6.5 \\
+\hspace*{0.2em}\color{gray}$|E|{=}165250$& 4&0.26&${\ge}1107$&\color{gray}\num{7.4e+06}&\color{gray}$>$ 1 h&\color{gray}\num{1.8e+09}&\color{gray}$>$ 12 h\\
+wing-nodal  & 2&0.01&29&648&3.5 &11580&2.9 \\
+\hspace*{0.2em}\color{gray}$|V|{=}10937$& 3&0.02&54&13039&15.2 &23709&17.9 \\
+\hspace*{0.2em}\color{gray}$|E|{=}75488$& 4&0.04&114&\num{6.0e+07}&1908.9 &113619&18.4 \\
     \bottomrule
@@
 
-\begin{figure}[t]
+\begin{figure}[h!] % try to keep this immediately above the next figure
     \centering
-    \input{gen-graph-omega}
+    \includegraphics{gen-graph-omega.pdf}
     \caption{Values of $\tilde{\omega}_k$ for random graphs $G(200, p)$, with varying edge probabilities.
@@
 
-\begin{figure}[t]
+\begin{figure}[h!] % try to keep this immediately below the previous figure
     \centering
-    \input{gen-graph-nodes}
+    \includegraphics{gen-graph-nodes.pdf}
     \caption{Search space size for random graphs $G(200, p)^k$, with varying edge probabilities. We see that
@@
 probability of random graphs having dominating vertices is very low.
+
+\begin{figure}[h!] % try to keep this immediately above the next figure
+    \centering
+    \includegraphics{gen-graph-omega-2.pdf}
+    \caption{The size of a maximum $2$-clique in random graphs $G(n, p)$ with varying edge
+        probabilities, and different values of $n$. For $G(50, p)$, a $2$-clique has size average 50
+    from $p = 0.42$ onwards.}
+    \label{figure:graph-omega-2}
+\end{figure}
+
+\begin{figure}[h!] % try to keep this immediately below the previous figure
+    \centering
+    \includegraphics{gen-graph-nodes-2.pdf}
+    \caption{The search space size for the maximum $2$-clique problem in random graphs $G(n, p)$
+        with varying edge probabilities, and for different values of $n$. As $n$ increases, the
+    complexity peak grows and moves slowly to the left.}
+    \label{figure:graph-nodes-2}
+\end{figure}
 
@@
 more pronounced, and shifts slightly towards the left (lower edge probabilities).
-
-\begin{figure}[t]
-    \centering
-    \input{gen-graph-omega-2}
-    \caption{The size of a maximum $2$-clique in random graphs $G(n, p)$ with varying edge
-        probabilities, and different values of $n$. For $G(50, p)$, a $2$-clique has size average 50
-    from $p = 0.42$ onwards.}
-    \label{figure:graph-omega-2}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-    \input{gen-graph-nodes-2}
-    \caption{The search space size for the maximum $2$-clique problem in random graphs $G(n, p)$
-        with varying edge probabilities, and for different values of $n$. As $n$ increases, the
-    complexity peak grows and moves slowly to the left.}
-    \label{figure:graph-nodes-2}
-\end{figure}
 

### 2016-07-01 talk/talk.tex
Start

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };

        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

\makeatletter
\newenvironment{nearlyplainframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{#2}
    }{
    \end{frame}
}
\makeatother

% }}}

\title[Finding Maximum $k$-Cliques Faster Using Lazy Global Domination]{Finding Maximum $k$-Cliques
Faster \\ Using Lazy Global Domination}
\author[Ciaran McCreesh and Patrick Prosser]{\textbf{Ciaran McCreesh} and Patrick
Prosser}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics*[keepaspectratio=true, width=\paperwidth]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Maximum Clique}
\end{frame}

\begin{frame}{Maximum $k$-Clique}
\end{frame}

\begin{frame}{Maximum $k$-Club}
\end{frame}

\begin{frame}{Existing Work}
    \begin{itemize}
        \item Many computational papers on the $k$-club problem
            \begin{itemize}
                \item Not a hereditary property, and modelling it is hard, so lots of scope for
                    being clever.
            \end{itemize}

        \item ``Unlike the maximum clique problem, the maximum $k$-clique problem has not been the
            subject of extensive research and we are not aware of any computational results for this
            problem to date.'' (Shahinpour and Butenko, 2013)
    \end{itemize}
\end{frame}

\begin{frame}{Reducing $k$-Clique to Clique}
\end{frame}

\begin{frame}{But Is This Practical?}
    \begin{itemize}
        \item We probably want to work with large sparse graphs as inputs.
        \item There are many good maximum clique algorithms for large sparse graphs.
        \item But $G^k$ is large and dense!
        \item There are also good maximum clique algorithms for small dense graphs, but small dense
            graphs can be \emph{very} hard.
            \begin{itemize}
                \item $\omega(\operatorname{G}(500, 0.9))$ can't be solved in a CPU-decade.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{The Good News}
    \begin{itemize}
        \item Start with a state-of-the-art maximum clique algorithm for small dense graphs.
        \item Replace cubic preprocessing and inference with cheaper quadratic algorithms.
        \item Spend some time making sure the $G^k$ reduction fast.
        \item Buy a hefty amount of RAM.
    \end{itemize}
\end{frame}

\begin{frame}{The Results}
    \begin{itemize}
        \item Erd\H{o}s collaboration graphs (up to 6,927 vertices, 11,850 edges)
            \begin{itemize}
                \item 16 take $<1s$, 2 take $<12s$, 3 $>1h$.
            \end{itemize}

        \item DIMACS 2 Clique Graphs (up to 500 vertices, 46,627 edges)
            \begin{itemize}
                \item 22 take $<1s$, 2 take $>1h$.
            \end{itemize}

        \item DIMACS 10 Partitioning Graphs (smallest 20, up to 36,519 vertices, 1,007,284 edges)
            \begin{itemize}
                \item 2 take $>1h$, most take well under a minute.
            \end{itemize}

        \item DIMACS 10 Clustering Graphs (smallest 20, up to 40,421 vertices, 175,691 edges)
            \begin{itemize}
                \item 7 take $>1h$, most take well under a minute.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{But We Can Do Better}
\end{frame}

\begin{frame}{Better Results!}
    \begin{itemize}
        \item Erd\H{o}s collaboration graphs
            \begin{itemize}
                \item The three open instances now take 3.8s, 69.4s, 207.3s.
            \end{itemize}

        \item DIMACS 2 Clique Graphs
            \begin{itemize}
                \item The two open instances now take 0.1s.
            \end{itemize}

        \item DIMACS 10 Partitioning Graphs
            \begin{itemize}
                \item One open instance now takes 16.9s, the other still takes $>1h$.
            \end{itemize}

        \item DIMACS 10 Clustering Graphs
            \begin{itemize}
                \item Two open instances closed.
            \end{itemize}

        \item Not a universal success, though
            \begin{itemize}
                \item Factor of 10 slowdown on a few fairly easy instances.
                \item Without laziness, the results are terrible\ldots
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{What About Parallel Search?}

    \begin{itemize}
        \item Active research topic for maximum clique.
            \begin{itemize}
                \item Work stealing strategies really make a huge difference (often more so than
                    load balance)!
                \item Use parallelism to steal early, where value ordering heuristics are
                    weakest---a bit like discrepancy search or restarts.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{Even Betterer Results!}
    \begin{itemize}
        \item Two open instances closed.
        \item Improved bounds on four remaining instances.
        \item Also super-linear speedups in (at least) two cases.
        \item But only if we get it right:
            \begin{itemize}
                \item Again, work stealing strategies matter.
                \item Load balancing is harder: must use a much deeper splitting limit than for
                    typical maximum clique graphs.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{What About Random Graphs?}
    \centering
    \input{gen-graph-nodes}
\end{frame}

\begin{frame}{Conclusion}

    \begin{itemize}
        \item There are many clique relaxations (density-based, degree-based, distance-based,
            \ldots). It's often not clear which you'd want in practice.
        \item $k$-Clique tends to be easy to calculate, at least.
        \item Usually the $k$-Clique and $k$-Club numbers are the same, too.
        \item The domination rule is probably useful in other settings.
        \item Work-stealing strategies matter when doing parallel search.
    \end{itemize}
\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{center}
        \vspace*{4em}
        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
        \vspace{1em}
        \href{mailto:c.mccreesh.1@research.gla.ac.uk}{\nolinkurl{c.mccreesh.1@research.gla.ac.uk}} \\
    \end{center}

    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };
    \end{tikzpicture}
\end{frame}

\end{document}

### 2016-07-01 talk/talk.tex
More talk

@@
 
-\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
+\tikzset{vertex/.style={draw, circle, inner sep=3pt, minimum size=0.5cm, font=\small\bfseries}}
 \tikzset{notvertex/.style={vertex, color=white, text=black}}
 \tikzset{plainvertex/.style={vertex}}
-\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
-\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
-\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
-\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
+\tikzset{selectedvertex/.style={vertex, fill=uofgcobalt}}
+\tikzset{selectedvertex2/.style={vertex, fill=uofglawn}}
+\tikzset{redvertex/.style={vertex, fill=uofgpillarbox}}
 \tikzset{edge/.style={color=black!50!white}}
@@
 \begin{frame}{Maximum Clique}
+    \centering
+    \begin{tikzpicture}
+        \newcount \c
+        \foreach \n in {1, ..., 8}{
+            \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+            \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 5 \OR \n = 8}{
+                \node[selectedvertex] (N\n) at (\the\c:2) {\n};
+            }{
+                \node[plainvertex] (N\n) at (\the\c:2) {\n};
+            }
+        }
+
+        \draw [edge] (N2) -- (N3);
+        \draw [edge] (N2) -- (N4);
+        \draw [edge] (N3) -- (N7);
+        \draw [edge] (N4) -- (N5);
+        \draw [edge] (N5) -- (N6);
+        \draw [edge] (N6) -- (N7);
+
+        \draw [bedge] (N1) -- (N2);
+        \draw [bedge] (N1) -- (N5);
+        \draw [bedge] (N1) -- (N8);
+        \draw [bedge] (N2) -- (N5);
+        \draw [bedge] (N2) -- (N8);
+        \draw [bedge] (N5) -- (N8);
+    \end{tikzpicture}
 \end{frame}
@@
 \begin{frame}{Maximum $k$-Clique}
+    \centering
+    \begin{tikzpicture}
+        \newcount \c
+        \foreach \n in {1, ..., 8}{
+            \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+            \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 3 \OR \n = 4 \OR \n = 5 \OR \n = 6 \OR \n = 8}{
+                \node[selectedvertex] (N\n) at (\the\c:2) {\n};
+            }{
+                \node[plainvertex] (N\n) at (\the\c:2) {\n};
+            }
+        }
+
+        \draw [edge] (N3) -- (N7);
+        \draw [edge] (N6) -- (N7);
+
+        \draw [bedge] (N1) -- (N2);
+        \draw [bedge] (N1) -- (N5);
+        \draw [bedge] (N1) -- (N8);
+        \draw [bedge] (N2) -- (N5);
+        \draw [bedge] (N2) -- (N8);
+        \draw [bedge] (N5) -- (N8);
+        \draw [bedge] (N2) -- (N3);
+        \draw [bedge] (N2) -- (N4);
+        \draw [bedge] (N4) -- (N5);
+        \draw [bedge] (N5) -- (N6);
+    \end{tikzpicture}
 \end{frame}
@@
 \begin{frame}{Maximum $k$-Club}
+    \centering
+    \begin{tikzpicture}
+        \newcount \c
+        \foreach \n in {1, ..., 8}{
+            \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+            \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 4 \OR \n = 5 \OR \n = 8}{
+                \node[selectedvertex] (N\n) at (\the\c:2) {\n};
+            }{
+                \ifthenelse{\n = 3 \OR \n = 6}{
+                    \node[redvertex] (N\n) at (\the\c:2) {\n};
+                }{
+                    \node[plainvertex] (N\n) at (\the\c:2) {\n};
+                }
+            }
+        }
+
+        \draw [edge] (N3) -- (N7);
+        \draw [edge] (N6) -- (N7);
+
+        \draw [bedge] (N1) -- (N2);
+        \draw [bedge] (N1) -- (N5);
+        \draw [bedge] (N1) -- (N8);
+        \draw [bedge] (N2) -- (N5);
+        \draw [bedge] (N2) -- (N8);
+        \draw [bedge] (N5) -- (N8);
+        \draw [bedge] (N2) -- (N3);
+        \draw [bedge] (N2) -- (N4);
+        \draw [bedge] (N4) -- (N5);
+        \draw [bedge] (N5) -- (N6);
+    \end{tikzpicture}
 \end{frame}
@@
 \begin{frame}{Reducing $k$-Clique to Clique}
+    \centering
+    \begin{tikzpicture}
+        \begin{scope}
+            \newcount \c
+            \foreach \n in {1, ..., 8}{
+                \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+                \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 3 \OR \n = 4 \OR \n = 5 \OR \n = 6 \OR \n = 8}{
+                    \node[selectedvertex] (N\n) at (\the\c:2) {\n};
+                }{
+                    \node[plainvertex] (N\n) at (\the\c:2) {\n};
+                }
+            }
+
+            \draw [edge] (N3) -- (N7);
+            \draw [edge] (N6) -- (N7);
+
+            \draw [bedge] (N1) -- (N2);
+            \draw [bedge] (N1) -- (N5);
+            \draw [bedge] (N1) -- (N8);
+            \draw [bedge] (N2) -- (N5);
+            \draw [bedge] (N2) -- (N8);
+            \draw [bedge] (N5) -- (N8);
+            \draw [bedge] (N2) -- (N3);
+            \draw [bedge] (N2) -- (N4);
+            \draw [bedge] (N4) -- (N5);
+            \draw [bedge] (N5) -- (N6);
+        \end{scope}
+        \begin{scope}[xshift=6cm]
+            \newcount \c
+            \foreach \n in {1, ..., 8}{
+                \c=\n \advance\c by -1 \multiply\c by -360 \divide\c by 8 \advance\c by 90 \advance\c by 22.5
+                \ifthenelse{\n = 1 \OR \n = 2 \OR \n = 3 \OR \n = 4 \OR \n = 5 \OR \n = 6 \OR \n = 8}{
+                    \node[selectedvertex] (N\n) at (\the\c:2) {\n};
+                }{
+                    \node[plainvertex] (N\n) at (\the\c:2) {\n};
+                }
+            }
+
+            \draw [edge] (N2) -- (N7);
+            \draw [edge] (N3) -- (N7);
+            \draw [edge] (N5) -- (N7);
+            \draw [edge] (N6) -- (N7);
+
+            \draw [bedge] (N1) -- (N2);
+            \draw [bedge] (N1) -- (N3);
+            \draw [bedge] (N1) -- (N4);
+            \draw [bedge] (N1) -- (N5);
+            \draw [bedge] (N1) -- (N6);
+            \draw [bedge] (N1) -- (N8);
+            \draw [bedge] (N2) -- (N3);
+            \draw [bedge] (N2) -- (N4);
+            \draw [bedge] (N2) -- (N5);
+            \draw [bedge] (N2) -- (N6);
+            \draw [bedge] (N2) -- (N8);
+            \draw [bedge] (N3) -- (N4);
+            \draw [bedge] (N3) -- (N5);
+            \draw [bedge] (N3) -- (N6);
+            \draw [bedge] (N3) -- (N8);
+            \draw [bedge] (N4) -- (N5);
+            \draw [bedge] (N4) -- (N6);
+            \draw [bedge] (N4) -- (N8);
+            \draw [bedge] (N5) -- (N6);
+            \draw [bedge] (N5) -- (N8);
+            \draw [bedge] (N6) -- (N8);
+        \end{scope}
+    \end{tikzpicture}
 \end{frame}
@@
 \begin{frame}{But We Can Do Better}
+    \centering
+    \begin{tikzpicture}
+        \begin{scope}
+            \node[selectedvertex] (Na) at (0, 2) {$a$};
+            \node[selectedvertex2] (Nb) at (6, 2) {$b$};
+
+            \node[plainvertex] (N1) at (3, 4) {$1$};
+            \node[plainvertex] (N2) at (3, 3) {$2$};
+            \node[plainvertex] (N3) at (3, 2) {$3$};
+            \node[plainvertex] (N4) at (3, 1) {$4$};
+            \node[plainvertex] (N5) at (3, 0) {$5$};
+
+            \draw [edge] (Na) -- (N1);
+            \draw [edge] (Na) -- (N2);
+            \draw [edge] (Na) -- (N3);
+            \draw [edge] (Na) -- (N4);
+            \draw [edge] (Na) -- (N5);
+
+            \draw [edge] (Nb) -- (N1);
+            \draw [edge] (Nb) -- (N3);
+            \draw [edge] (Nb) -- (N5);
+        \end{scope}
+    \end{tikzpicture}
 \end{frame}

### 2016-07-01 talk/talk.tex
Punctuation

@@
     \begin{itemize}
-        \item Many computational papers on the $k$-club problem
+        \item Many computational papers on the $k$-club problem.
             \begin{itemize}
@@
     \begin{itemize}
-        \item Erd\H{o}s collaboration graphs (up to 6,927 vertices, 11,850 edges)
+        \item Erd\H{o}s collaboration graphs (up to 6,927 vertices, 11,850 edges):
             \begin{itemize}
@@
 
-        \item DIMACS 2 Clique Graphs (up to 500 vertices, 46,627 edges)
+        \item DIMACS 2 Clique Graphs (up to 500 vertices, 46,627 edges):
             \begin{itemize}
@@
 
-        \item DIMACS 10 Partitioning Graphs (smallest 20, up to 36,519 vertices, 1,007,284 edges)
+        \item DIMACS 10 Partitioning Graphs (smallest 20, up to 36,519 vertices, 1,007,284 edges):
             \begin{itemize}
@@
 
-        \item DIMACS 10 Clustering Graphs (smallest 20, up to 40,421 vertices, 175,691 edges)
+        \item DIMACS 10 Clustering Graphs (smallest 20, up to 40,421 vertices, 175,691 edges):
             \begin{itemize}
@@
     \begin{itemize}
-        \item Erd\H{o}s collaboration graphs
+        \item Erd\H{o}s collaboration graphs:
             \begin{itemize}
@@
 
-        \item DIMACS 2 Clique Graphs
+        \item DIMACS 2 Clique Graphs:
             \begin{itemize}
@@
 
-        \item DIMACS 10 Partitioning Graphs
+        \item DIMACS 10 Partitioning Graphs:
             \begin{itemize}
@@
 
-        \item DIMACS 10 Clustering Graphs
+        \item DIMACS 10 Clustering Graphs:
             \begin{itemize}
@@
 
-        \item Not a universal success, though
+        \item Not a universal success, though:
             \begin{itemize}

### 2016-07-03 talk/talk.tex
Formatting

@@
 
-\begin{frame}{The Results}
+\begin{frame}{The Results ($k \in \{2, 3, 4\}$)}
     \begin{itemize}
-        \item Erd\H{o}s collaboration graphs (up to 6,927 vertices, 11,850 edges):
+        \item Erd\H{o}s collaboration graphs:
             \begin{itemize}
+                \item $|V| \le 6,927$, $|E| \le 11,850$, $D(G^k) \le 0.57$ (or $= 1$).
                 \item 16 take $<1s$, 2 take $<12s$, 3 $>1h$.
@@
 
-        \item DIMACS 2 Clique Graphs (up to 500 vertices, 46,627 edges):
+        \item DIMACS 2 Clique Graphs with diameter $> 2$:
             \begin{itemize}
+                \item $|V| \le 500$, $|E| \le 46,627$, $D(G^k) \le 0.87$ (or $= 1$)
                 \item 22 take $<1s$, 2 take $>1h$.
@@
 
-        \item DIMACS 10 Partitioning Graphs (smallest 20, up to 36,519 vertices, 1,007,284 edges):
+        \item DIMACS 10 Partitioning Graphs (smallest 20):
             \begin{itemize}
+                \item $|V| \le 36,519$, $|E| \le 1,007,284$, $D(G^k) \le 0.67$.
                 \item 2 take $>1h$, most take well under a minute.
@@
 
-        \item DIMACS 10 Clustering Graphs (smallest 20, up to 40,421 vertices, 175,691 edges):
+        \item DIMACS 10 Clustering Graphs (smallest 20):
             \begin{itemize}
+                \item $|V| \le 40,421$, $|E| \le 175,691$, $D(G^k) \le 0.99$.
                 \item 7 take $>1h$, most take well under a minute.

### 2016-07-08 talk/talk.tex
Extra line
More on work stealing
Typo
Some tweaks

@@
             \end{itemize}
+        \item It's well known in the CP community that reduction to clique is a bad way of solving
+            things.
+            \begin{itemize}
+                \item (This is not necessarily true, but it's well known\ldots)
+            \end{itemize}
     \end{itemize}
@@
         \item Replace cubic preprocessing and inference with cheaper quadratic algorithms.
-        \item Spend some time making sure the $G^k$ reduction fast.
+        \item Spend some time making the $G^k$ reduction fast.
         \item Buy a hefty amount of RAM.
@@
             \draw [edge] (Nb) -- (N5);
+
+            \draw [edge, dotted] (Na) to [out=-15, in=195] (Nb);
         \end{scope}
@@
 \begin{frame}{What About Parallel Search?}
-
     \begin{itemize}
@@
                     load balance)!
+            \end{itemize}
+        \item In this paper:
+            \begin{itemize}
+                \item Dynamic work splitting, starting at the top, and working downwards.
                 \item Use parallelism to steal early, where value ordering heuristics are
@@
 
+\begin{frame}{Slide of Blatant Propaganda}
+    \begin{itemize}
+        \item This technique is the only known practical way of getting:
+            \begin{itemize}
+                \item Reproducible runtimes.
+                \item Guarantees of no (exponential) slowdown over sequential, or when adding cores.
+                \item Decent work distribution.
+                \item Respectable speedups in practice.
+            \end{itemize}
+        \item I personally think these properties are all critical for practical multi-core parallel
+            search.
+    \end{itemize}
+\end{frame}
+
 \begin{frame}{What About Random Graphs?}
@@
 
+\begin{frame}{Conspicuously Missing From This Paper}
+    \begin{itemize}
+        \item Parallel $G^k$ construction.
+        \item Predicting $G(n, p)^k$ on random graphs.
+        \item So just why are $G^k$ graphs so easy to solve, anyway?
+    \end{itemize}
+\end{frame}
+
 \begin{frame}{Conclusion}
-
     \begin{itemize}
